### 2016-06-23 notes/Marillia/Jun23-2016/notes.tex
added Peel's code

\documentclass[12pt]{article}
\usepackage[top=0.85in, bottom=0.93in, right=0.93in, left=0.93in,
paperwidth=8.5in, paperheight=11in, nohead]{geometry}
\geometry{letterpaper}
\usepackage[pdftex]{graphicx}
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{setspace}
\usepackage{mdwlist}
\usepackage{verbatim}
\usepackage[numbers]{natbib}

%% make bibliography more compact \setlength{\bibsep}{0in}
\renewcommand\bibsection{\subsubsection*{\refname}}

\begin{document}

\begin{centering}
  \large {\bf The assembly of plant-pollinator communities in response
  to restoration8} \\
\end{centering}
\vspace{0.15in}

\section{Questions}
\begin{enumerate}
\item Determine whether there are change points in the assmebly of the
  ``maturing'' hedgerow communties, and compare those to the
  ``mature'' and ``unrestored''
\item Examine the resileicen of the maturing communities through time,
  as well as the different site types
\item Simulate the extinction of plants affected by drought, and see
  if the predict extinct pollinator species that are 
\item Assess whether (and how) species roles/importance are changing
  through time, as community matures. 
\end{enumerate}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End: