### 2016-11-02 ms/coverLetter/coverLetter.tex
added cover letter

\documentclass[12pt]{letter}
\usepackage[margin=2.5cm]{geometry}
\usepackage{mathpazo}
\usepackage[pdftex]{graphicx}
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{times}
\usepackage{textcomp}

\address{Dept of Environmental Science,\\
  Policy, and Management\\
  130 Mulford Hall\\
  University of California, Berkeley\\
  Berkeley, California, 94720, USA} \raggedright
\begin{document}

\begin{letter}{}

  \opening{Dear editor,}

  Please find attached the manuscript \textit{Major interaction
    reorganizations punctuate the assembly of pollination networks}
  which we are submitting for consideration as a Letter in
  \textit{Ecology Letters}.

  When a species goes extinct, more than a species is lost. That
  species interacted with other species in its community, likely in
  ways that were important or even crucial to others' survival. As the
  world continues to lose species at an alarming rate, it has become
  increasingly imperative to aid the recovery of lost interactions and
  component biodiversity through ecological restoration. We know
  little, however, about how to re-assemble interacting communities
  through restoration, or the process of ecological network assembly
  more generally.

  Using extensive surveys of pollinators spanning eight years
  comprising \texttildelow $20,000$ pollinator visitation records, we
  explore the assembly of plant-pollinator communities at native plant
  restorations in the Central Valley of California. Employing newly
  developed methods for examining temporal changes in networks, we
  find that species are highly dynamic in their network position,
  causing community assembly to be punctuated by major interaction
  reorganizations. The most persistent and generalized species were
  also the most variable in their network positions, contrary to what
  is expected through preferential attachment theory --- an assembly
  theory otherwise well-supported in the network literature. Our study
  is the first long-term study on the temporal assembly of ecological
  networks. It also challenges the hypothesis that mutualistic systems
  assemble through preferential attachment (Bascompte and Stouffer,
  2009).

  Our manuscript is original and was carried out fully by the authors.
  All authors agree with the contents of the manuscript.  This
  manuscript is not published, nor is it in consideration for
  publication elsewhere.  All research not of the authors' is fully
  acknowledged.  The authors declare no conflict of interest. All
  appropriate ethical standards were followed.  Thank you for
  reviewing our manuscript and we hope you will find it suitable for
  publication.

  Regards,
  Lauren C. Ponisio, PhD\\
  Claire Kremen, Professor

\end{letter}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End: