### 2016-11-07 ms/coverLetter/coverLetter.tex
minor changes to cover letter

\documentclass[12pt]{letter}
\usepackage[margin=2.5cm]{geometry}
\usepackage{mathpazo}
\usepackage[pdftex]{graphicx}
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{times}
\usepackage{textcomp}

\address{Dept of Environmental Science,\\
  Policy, and Management\\
  130 Mulford Hall\\
  University of California, Berkeley\\
  Berkeley, California, 94720, USA} \raggedright
\begin{document}

\begin{letter}{}

  \opening{Dear editor,}

  W are pleased to submit our manuscript \textit{Major interaction
    reorganizations punctuate the assembly of pollination networks}
  for consideration as a Letter in \textit{Ecology Letters}.

  Our research deals with two fundamental aspects of ecological
  theory: unerstanding how species-rich communities assemble, and how
  these assemblages change through time. As the world continues to
  lose species at an alarming rate, it has become increasingly
  imperative to aid the recovery of lost interactions and component
  biodiversity through ecological restoration. When a species goes
  extinct, not only a species is lost, but also its interactions. We
  know little, however, about how to re-assemble interacting
  communities through restoration, or the process of ecological
  network assembly more generally.

  Using extensive surveys of pollinators spanning eight years
  comprising \texttildelow $20,000$ pollinator visitation records, we
  explore the assembly of plant-pollinator communities at native plant
  restorations in the Central Valley of California. For the first time
  in the ecological litterature we employ a newly developed method to
  examine the temporal changes in networks. Among other things, we
  find that species are highly dynamic in their network position,
  causing community assembly to be punctuated by major interaction
  reorganizations. The most persistent and generalized species are
  also the most variable in their network positions, contrary to what
  is expected through preferential attachment theory --- an assembly
  theory otherwise well-supported in the network literature. Our study
  is the first long-term study on the temporal assembly of ecological
  networks.  Our results are compelling and provide empirical evidence
  that widen our understading on how communities assembly and how
  species interactions changes through time. Furthermore, our results
  also contribute to the knowledge of how communities will be able to
  maintain function in the face of species extinction. And finally,
  our results challenge the view that communities assemble through
  preferential attachment. We believe that these exciting results that
  link three major ecological fields (interaction networks, community
  dinamics and restauration ecology) that will be of broad interest to
  the readership of \textit{Ecology Letters}.

  % Our manuscript is original and was carried out fully by the authors.
  % All authors agree with the contents of the manuscript.  This
  % manuscript is not published, nor is it in consideration for
  % publication elsewhere.  All research not of the authors' is fully
  % acknowledged.  The authors declare no conflict of interest. All
  % appropriate ethical standards were followed. 
  Thank you for reviewing our manuscript and we hope you will find it
  suitable for publication.

  Regards,
  Lauren C. Ponisio, PhD\\
  Claire Kremen, Professor
%is there a reason why I am not here? Just curious =0) 
\end{letter}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End: