### 2016-12-01 ms/ms.tex
updated cover letter, ms

\documentclass[12pt]{article} 
\usepackage[pdftex]{graphicx}
\usepackage[numbers,sort&compress]{natbib} 
\usepackage{color}
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{verbatim}
\usepackage{mathpazo} 
\usepackage{setspace}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{lscape}
\usepackage{fancyhdr}
\usepackage[normalem]{ulem} 
\usepackage{hyperref}
\usepackage[parfill]{parskip}
\usepackage{xr}
\externaldocument{SI}
\usepackage{url}
\usepackage{rotating}

\hypersetup{colorlinks=true, linkcolor=black, citecolor=black}
\RequirePackage{lineno}

\newcommand{\flagged}[1] {
  \textcolor{blue}{#1}
}

\def\title{Major network reorganizations punctuate the assembly of
  plant-pollinator communities}

\def\author{Lauren C.\
  Ponisio$^{1,2,3}$, Marilia P. Gaiarsa$^4$, Claire Kremen$^1$}

\def\runninghead{Plant-pollinator network assembly}
\def\keywords{community assembly, change points, robustness,
  nestedness, modularity, bipartite, preferential attachment}

\def\extras{
  \begin{itemize}
  \item Submitted as a Letter
  \item Abstract word count: 148
  \item Main text word count: 4967
  \item Number of references: 65
  \item Number of figures: 6
  \item Author contributions: CK designed the study, LCP and CK
    collected data, LCP and MPG analyzed output data.  LCP wrote the
    first draft of the manuscript, and all authors contributed to
    revisions.
  \item Data accessibility statement: Data will de depositied in Dryad
    upon publication
  \end{itemize}
}

\def\affiliation{
  \begin{enumerate}
  \item Department of Environmental Science, Policy, and Management\\
    University of California, Berkeley\\
    130 Mulford Hall\\
    Berkeley, California, USA\\
    94720\\
  \item Berkeley Institute for Data Science (BIDS)\\
    University of California, Berkeley\\
    190 Doe Library \\
    Berkeley, California, USA\\
  \item Department of Entomology\\
    University of California, Riverside\\
    417 Entomology Bldg.\\
    Riverside, California, USA\\
    92521\\
  \item Departamento de Ecologia\\
    Universidade de S{\~a}o Paulo\\
    S{\~a}o Paulo, SP, Brazil\\
    05508-900\\
  \end{enumerate}
}

\newcommand{\mstitlepage}{
  \paragraph{Running head:} \textsc{\runninghead}
  % \parindent=0pt
  \begin{center}%
    {\LARGE \title \par}%
    \vskip 3em%
    {\large
      \lineskip .75em%
      \begin{tabular}[t]{c}%
        \author
      \end{tabular}\par}%
    \vskip 1.5em%
  \end{center}\par
  \affiliation
}
\clearpage

\begin{document}

\mstitlepage
\doublespacing
\linenumbers
\clearpage

\begin{abstract}  
  As the world continues to lose species at an alarming rate,
  anticipating a community's ability to resist collapse will depend on
  the strength of interaction networks. We know little, however, about
  how to re-assemble interacting communities through restoration, or
  the process of ecological network assembly more generally. Using an
  eight-year dataset comprising nearly $20,000$ pollinator visitation
  records, we explore the assembly of plant-pollinator communities at
  native plant restorations sites in an agricultural landscape. We
  find that species occupy highly dynamic network positions through
  time, causing the community assembly process to be punctuated by
  major network reorganizations. In contrast, the non-assembling
  networks do not restructure as frequently. The most persistent and
  generalized species are also the most variable in their network
  positions, contrary to what is expected by assembly following
  preferential attachment --- a widly supported assembly
  mechanism. Our results fundamentally alter our understanding of how
  communities assembly and how species interactions change through
  time while helping to inform efforts to re-assembly robust
  communities through restoration.
\end{abstract}

\keywords

\clearpage

\section*{Introduction}
\label{sec:introduction}

%% introduce mutualisms, restoration, network assembly
When a species goes extinct, more than a species is lost. That species
interacted with other species in its community, likely in ways that
were important or even crucial to others' survival. When a community
--- a unique combination of interacting species --- loses one species,
the system becomes weaker. Global change has created a severe
biodiversity crisis, and as species are lost, so are their
interactions \citep{dunn2009sixth, barnosky2011has}. A community's
ability to resist collapse will depend on the strength of each
interaction network \citep{Memmott2004, Rezende2007, Bascompte2009,
  Thebault2010}. To safeguard ecological function, it has become
increasingly imperative to aid the recovery of lost interactions and
component biodiversity and by facilitating the assembly of robust
interaction networks via ecological restoration
\citep{menz-2010-4}. However, we know very little about how ecological
networks assemble in general, rendering restoration of interacting
communities quite difficult.

%% most people think network assemble via preferential attachment, and
%% there is some evidence of that. (other studies on mutualisms?)

Preferential attachment, the most widely explored mechanism of network
assembly, predicts that species entering a network are more likely to
interact with species that are already well-connected \citep[the
``rich-get-richer'' principle,][]{barabasi1999emergence}. In
pollination systems --- a particularly ubiquitous mutualism
\citep{ollerton-2011-321, klein-2007-303} --- some studies have found
support for this assembly mechanism. Investigating the day-to-day,
temporal assembly of a plant-pollinator network within a season,
\cite{Olesen2008} found that phenologically new plant and pollinator
species tended to interact with already well-connected species,
potentially because these species are either more abundant or
temporally persistent. Using a space-for-time substitution to study
primary succession, \cite{albrecht2010plant} also found evidence that
assembly along a glacier foreland occurred through preferential
attachment. Specifically, network nestedness (i.e, a core group of
generalists interacts with both specialist and generalist species)
increased as the community aged \citep{albrecht2010plant}. An increase
in nestedness could have occurred via the preferential attachment
process, whereby specialist species attach to the well-connected,
generalist core.

Preferential attachment, the most widely explored mechanism of network
assembly, \citep{barabasi1999emergence}, predicts that species
entering a network are more likely to interact with species that are
already well-connected \citep[the ``rich-get-richer''
principle,][]{barabasi1999emergence}. In pollination systems --- a
particularly ubiquitous mutualism \citep{ollerton-2011-321,
  klein-2007-303} --- some studies have found support for this
assembly mechanism. Investigating the day-to-day, temporal assembly of
a plant-pollinator network within a season, \cite{Olesen2008} found
that phenologically new plant and pollinator species tended to
interact with already well-connected species, potentially because
these species are either more abundant or more temporally
persistent. In addition, using a space-for-time substitution to study
primary succession along a glacier foreland, \cite{albrecht2010plant}
also found some evidence that assembly occurred through preferential
attachment. Specifically, network nestedness (i.e, a core group of
generalists interacts with both specialist and generalist species)
increased as the community aged \citep{albrecht2010plant}. An
increase in nestedness could result from the preferential attachment
process where specialist species attach to the well-connected,
generalist core.

%% there are other ways to assemble that are not preferentia;
%% attchment, particularly vis big reorganizations of interactions
%% Lauren I love this paragraph!
In contrast to the network build-up described by preferential
attachment, significant reorganizations of interactions can also
punctuate assembly \citep{peel2014detecting}. Such significant
reorganizations of interactions, or network changing points, are
observed in social networks that respond to abrupt shifts in the
behavior of interactors \citep{peel2014detecting}. A change point is
caused by a merge, split, fragmentation or formation of modules within
a network. In ecological communities, such shifts may occur if, as new
species colonize, resident species change their interaction partners
to optimize their foraging effort. In plant-pollinator communities,
theory predicts that pollinators optimize their use of floral
resources to reduce interspecific competition and improve resource-use
efficiency \citep{pyke1984optimal, valdovinos2010consequences,
  valdovinos2013adaptive, albrecht2010plant, Bluthgen2007}. No
studies, however, have examined if network changing points occur
during ecological network assembly and/or how these changes relate to
species behavior.

%% restoration
Understanding network assembly is particularly relevant to ecological
restoration, which is essentially ``applied succession''
\citep[e.g.,][]{parker1997scale}.  In pollination systems, time has
been shown to affect the structure of networks in restored areas
\citep{forup-2008-742, devoto2012understanding}, suggesting that
interactions change as a community develops. Facilitating network
restoration is especially imperative in areas where species
interactions provide essential ecosystem services, such as crop
pollination in agricultural landscapes. To promote pollinator services
in agriculture, some farmers plant strips of native plants along farm
edges (hedgerows).  By providing habitat, hedgerows augment the
richness, abundance and trait diversity of pollinators in agricultural
landscapes \citep{morandin-2013-829, kremen-2015-602,
  ponisio2015farm}, and promote the persistence and colonization of
floral resource specialists \citep{mgonigle-2015-x}. As the community
assembles, it is important to understand how these new species are
incorporated into the network as well as the consequences of adding
species for interaction patterns and robustness.

%% Lauren I would put 10 year dataset; and I *loved* the way you
%% organized the ideas here!!
We explore the process of network development using a nine-year
dataset of plant-pollinator community assembly following hedgerow
restoration in the highly simplified and intensively managed
agricultural landscape of California’s Central Valley. We first
determine whether the mechanism underlying network assembly represents
a buildup of interactions as would be predicted by preferential
attachment or significant reorganizations of interactions (i.e.,
network changing points). Even with changing points in interaction
organization, networks could still assemble via preferential
attachment if the network reorganizations were primarily driven by
peripheral, temporally variable species while a stable, well-connected
core of species persist. Therefore, we tested whether the species that
are most variable in their network position --- and thus important
contributors to network reorganizations -- are less persistent and
connected species. To further explore the mechanisms underlying the
temporal dynamics of the networks, also we examine patterns in the
species and interaction temporal turnover. Finally, we investigate
whether networks assemble toward predictable interaction patterns, and
the ramifications for the robustness of the networks.

%% objectives
% 1) Are hedgerows assembling by preferential attachment, or are there
% more abrupt reorganizations of network structure?
% 2) Who are the species responsible for interaction reorganizations?
% how to incorporate this? 
% 3) Are the changing points a product of turnover in species
% composition, or are more substantial changes in the verticles of the
% networks?
% 4) Is the structure of interactions changing? Does this change the
% robustness of the network to species extinction and/or perturbation?

\section*{Materials \& Methods}
\label{sec:methods}

\subsection*{Study sites and collection methods}
\label{sec:study-sites}

Study sites were located in the Central Valley of California in Yolo,
Colusa and Solano Counties. This area is composed of intensively
managed agriculture --- primarily monocultures of conventional row
crops, vineyards and orchards. Hedgerows border large
(ca. $30$-hectare) crop fields and measured between $3-6$ m wide and
approximately $350$ m long. Hedgerows consist of native, perennial,
shrub and tree plantings \citep[Fig.~S1, ][]{menz-2010-4,
  kremen-2015-602, mgonigle-2015-x}.

We selected five farm edges to be restored as hedgerows. We monitored
those locations before the hedgerow was planted, and tracked the
assembly of the community for up to seven years as the hedgerow
matured (from $2006-2014$). For controls, we concurrently monitored
non-assembling communities within mature hedgerows (greater than $10$
years since planting, $N=29$) and unrestored, weedy field margins
($N=19$). The mean distance between monitoring sites was $15$ km, and
the minimum distance between sites sampled in the same year was $1$
km. The entire area surveyed spanned almost $300$ $km^2$. The crop
fields adjacent to all sites were similarly managed as intensive,
high-input monoculture.

Sites were sampled between two and five times per year (Tables S1-S3,
mean $3.4$ samples per year). In each round of sampling, the order in
which sites were sampled was randomized. Surveys were conducted under
sunny conditions when the temperature was above $21^{\circ}\mathrm{C}$
and wind speed was below $2.5$ meters/second.

During each sampling round, flower-visitors to plants in assembling
hedgerows and non-assembling controls were netted for one hour of
active search time (the timer was paused when handling specimens). The
identify of the plant being visited was recorded for each floral
visitor. Honeybees (\textit{Apis mellifera}) were not collected
because their abundance is determined largely by hive placement by
bee-keepers. All other insect flower visitors that touched the
reproductive parts of the flower were collected; however, here we
focus only on wild bees and syrphids --- the most abundant and
effective pollinators in the system (representing 49 and 19 percent of
records, respectively, C.~Kremen, A.~Klein and L.~Morandin,
unpublished data). Bee and syrphid specimens were identified to
species (or morpho-species for some bee specimens in the genera
\textit{Nomada} and \textit{Sphecodes}) by expert taxonomists.

Quantitative networks were generated for each site through
time. Because the nuber of sampling rounds varied between years
(Tables S1-S3), we used the mean of the interactions between a pair of
plants and pollinators within a year to represent interaction
frequency.

\subsection*{Change point analysis}
\subsubsection*{Identifying change points}
We employed a change point detection method \citep{peel2014detecting}
to identify fundamental reorganizations in large-scale interaction
patterns. Change point detection methods have yet to be generalized to
quantitative networks, so for this analysis we focused on qualitative
(binary) networks. Following \cite{peel2014detecting}, we first
defined a probability distribution over the networks using the
generalized hierarchical random graph model (GHRG). The GHRG model
captures both assortative and disassortative structure patterns at all
scales in the network \citep{peel2014detecting}. A network $G$ is
composed of vertices $V$ and edges $E$. The GHRG model decomposes the
$N$ vertices into a series of nested groups, the relationships among
which are represented by the dendrogram $T$. The tips of $T$ are the
vertices of $G$, and the probability that two vertices $u$ and $v$
connect is given by the parameter $p_r$. The probability distribution
of the network $G$ is thus defined as:

\begin{equation}
  \label{eq:lik}
  P(G|T,{pr}) = p_r^{E_r}(1-p_r)^{N_r-E_r}
\end{equation}
% 
Where $E_r$ is the observed number of edges between vertices with the
common ancestor $r$, and $N_r$ is the total possible edges, and the
rest of the notation is described above.

Using Bayesian posterior inference and techniques from phylogenetic
tree reconstruction, we fit the GHRG model to the networks
\citep{peel2014detecting}. This is accomplished by using a Markov
chain Monte Carlo (MCMC) procedure to first sample the posterior
distribution of bipartitions, from which a consensus tree is derived
\citep{peel2014detecting}. We use $\beta$ distributions with the
hyperparameters $\alpha=\beta=1$ to define priors for $p_r$.

Once the GHRG model has been fit to the networks, we determine whether
a change point occurred between two time slices. To detect a change
point, we use Bayes factors to compare the fit of two models --- one
where a change point occurred between two networks, and one where no
change occurred. We chose a sliding window of length, $w$, of four
years to detect change points. Larger windows allow for more gradual
changes, and four was the maximum possible with our eight years of
data. Lastly, to calculate a $p$-value for the Bayes factors, we use
parametric bootstrapping to numerically estimate the null distribution
\citep{peel2014detecting}. We employed code published online by
L.~Peel for the change point analysis. Analyses were conducted in
Python 3.4.

We next tested whether the change points occurring in maturing
hedgerows were a component of the assembly process or a product of
environmental shifts that lead to network reorganizations in all types
of communities. We modeled the number of change points as successes
and the total number of sample years at each site as trials, and used
a generalized linear model with Binomial error to test whether the
probability of a change point occurrence varied by site type. We used
standard techniques to determine whether the assumptions of the models
were met for this and all subsequent models. For the non-assembling
controls, only sites with five or greater survey years were included
in this analysis ($N=11$). All statistical analyses were conducted in
R 3.2.3 \citep{R}.

\subsubsection*{Characteristics of species that contribute to change
  points}

To further elucidate the nature of the change points, we examine the
characteristics of the species that contributed to interaction
reorganization. Some species remained in relatively similar network
positions through time, whereas others were more variable in their
position and thus contributed more strongly to network
reorganization. We tested whether the more persistent species with the
highest degree (number of different interaction partners) were the
most stable in their network positions, as would be expected if the
networks were assembling via preferential attachment.

We calculated species persistence as the proportion of surveys in
which a plant or pollinator is observed. Species observed consistently
within and between years were thus maximally persistent. Weighted
species degree is calculated from interaction observations from an
extensive dataset from Yolo County (approx.~18000 interaction records)
that included both the data included in this study and additional data
from sites where we collected flower visitors using the same methods
\citep{mgonigle-2015-x, ponisio2015farm}. To represent network
position variability, we computed the coefficient of variation of
weighted closeness centrality \citep{freeman1978centrality} at each
site through time. Closeness centrality represents the importance of a
space by calcuating the path lengths to other vertices (species) in
the network \citep{freeman1978centrality}. The shorter the mean path
length to other species, the higher the closeness centrality. We used
linear mixed models to test whether the species closeness variability
(log) is related to the persistence or degree of that species
\citep{lme4, lmetest}. We included random effects for species and
site. Because the degree and persistence of pollinators were strongly
correlated, ($\rho = 0.071$, $p$-value $<$ $2*10^{-16}$), we included
each explanatory variable in separate models. Plant degree and
persistence were not significantly correlated, but we used the same
models as we did for the pollinators for consistency.  Because an
approximately logarithmic increase in closeness centrality --- as
would be expected with assembly by preferential attachment --- we also
tested whether log closeness centrality increases through time.

\subsection*{Species and interaction turnover}

%%Lauren I really liked the way you explained this part =]
Reorganizations of network structure can be the result of species
turnover or species changing their interaction partners (i.e.,
re-wiring). To better understand the mechanisms underlying the
temporal dynamics of the assembling networks, we examined patterns of
species and interaction turnover. For example, assembling networks may
have higher rates of pollinator turnover than non-assembling
communities because new pollinator species are colonizing and
establishing themselves \citep{mgonigle-2015-x}. Similarly, because
species are turning over and pollinators are trying to maximize their
foraging efficiency based on the species present, interactions may
turnover more quickly than in established communities. In addition, at
assembling hedgerows, plants that are unvisited in early years may
appear to ``colonize'' the networks as they became more attractive
resources and establish new interactions with pollinators.

To estimate the temporal species and interaction turnover, we used an
approach similar to calculating spatial $\beta$-diversity. Instead of
calculating the variation in community composition across sites within
a year, we estimated turnover across years at a site. We first
calculated the pairwise dissimilarity of plants, pollinators and
interactions between years within each site using the Chao
dissimilarity estimator that incorporates abundances, while also
accounting for unobserved records \citep{chao-2005-148}. Dissimilarity
estimates can be affected by the total number of species and
individuals sampled at a site \citep[e.g.,][]{ponisio2015farm}. For
example, the probability that two sites do not share any species is
higher when there are few individuals at those sites. Following
\cite{ponisio2015farm}, we used null models that constrained species
richness to estimate the deviation of the observed dissimilarity from
that which would be expected under a random community assembly
process. With the corrected dissimilarity values, we then calculated
the multivariate dispersion of community composition across years
\citep{anderson-2011-19}. In order to test whether assembling
hedgerows had more species and interactions turnover than
non-assembling communities, the species and interaction temporal
turnover estimates were modeled as responses in a linear mixed model
with site type as an explanatory variable and site as a random effect
\citep{lme4, lmetest}.

Though species may turnover across years, some groups of species may
essentially replace each other if they fill similar roles in the
network, occupying the same network position and interacting with
similar species. At non-assembling communities, species turnover may
overestimate the temporal changes in the networks if the interactions
occurring in one year are similar to those in the next year when they
are weighted by the similarity of their constituent species
(Fig.~\ref{fig:methods}). We developed a method to examine the
temporal turnover of interactions with weightings based on their
similarity. We followed the algorithm of \cite{ahn2010link} to cluster
all the interactions (edges) hierarchically across sites and years
based on their similarity, and build a dendrogram. The interaction
similarity is based how many plants and pollinators (vertices) two
edges share \citep{ahn2010link, kalinka2011linkcomm}. The more species
edges shared in common, the shorter the branch length between them on
the dendrogram.  We next calculated the temporal turnover of
interactions weighted by their similarity, as approximated by
``phylogenetic'' distance \citep{graham2008phylogenetic,
  picante-2010-1463}. We then used linear mixed models to test whether
the weighted turnover of interactions varied between assembling and
non-assembling networks, including site as a random effect
\citep{lme4, lmetest}.

\section*{Temporal changes in interaction patterns}

\subsection*{Network structure}
Any changing points in network structure may contribute to the
reorganization of the assembling networks into predictable interaction
patterns. Pollination networks exhibit two main structural patterns
--- modularity \citep[e.g.,][]{Olesen2007} and nestedness
\citep[e.g.,][]{Bascompte2006, Bascompte2003}. In modular networks,
interactions are insular, occurring within separate groups or
``modules'' more often than between modules. Modules in the network
may fragment as the network assembles, enhancing
modularity. Conversely, nested networks are like a pyramid of
interactions, where there are some species that interact with many
species, other species that interact with a subset of those species,
and so on. If species entering the network tend to interact with the
generalist base of the network pyramid as would be expected with
preferential attachment, nestedness would increase through time. The
connectance --- the proportion of observed out of possible
interactions --- would decrease as new, specialist species,
preferentially attach to the core. Finally, the overall level of
network specialization may change as the community
assembles. Network-level specialization will increase if specialist
species colonize the network or species begin to limit their
interaction niche breath as the network assembles
\citep{bluthgen-2006-9}.

To evaluate network nestedness, we used the estimator weighted NODF
\citep{almeida-neto-2008-1227}. NODF evaluates whether species with
fewer partners interact with subsets of partners with which more
connected species interact \citep{almeida-neto-2008-1227}. To estimate
modularity, we used a hierarchical clustering algorithm
\citep{Newman2004, csardi-2006}. We evaluate network specialization with
the metric H2, which estimates the deviation of the observed
interaction frequency between plants and pollinators from a null
expectation where all partners interact in proportion to their
abundances \citep{bluthgen-2006-9}. It ranges from zero for
generalized networks to one for specialized networks.  We calculated
standardized $z$-scores so that nestedness, modularity and
specialization metrics could be compared across communities. The
$z$-scores were calculated by generating an ensemble of $999$ randomly
assembled communities, subtracting the mean of the statistic
calculated across these communities from the observed value, and then
dividing by the standard deviation. To assemble random communities, we
reshuffled the interactions between species but fixed the total number
of interactions, species and interaction frequency distributions
\citep{Galeano2009}.

To test whether network modularity, nestedness, connectance or
specialization changed linearly with assembly, we usedd linear mixed
models with the descriptive network metrics as the response variable,
year of assembly as the explanatory variable, and random effects of
site and year. The number of species in a network affects the patterns
of interaction possible, so we also examined the change in plant and
pollinator species richness through time. We employ generalized linear
mixed models with Poisson error to model richness \citep{lme4}. We
scaled explanatory variables.

\subsection*{Network robustness}
Lastly, we tested whether the changes in interaction patterns
associated with network assembly affect the robustness of the network
to species loss and cascading perturbations. Following
\cite{Memmott2004}, we simulated plant species extinction and the
subsequent extinction cascades of pollinator species. Becaused the
reproduction of plant species is facilitated by active restoration
efforts, it is unlikely the extinction of pollinator species would
affect plant populations in the hedgerows. However, plants ceasing to
bloom, for example in response to drought, will likely affect the
pollinators that depend on them. We eliminated plants species based on
their degree or abundance, and then calculated the number of
pollinators that secondarily went extinct. The area below the
extinction curve is an estimate of network robustness
\citep{Memmott2004, bipartite}.

We also explored how the robustness to cascading perturbations changed
as the community assembled, using algebraic connectivity --- the
second smallest eigenvalue of the Laplacian matrix
\citep{fiedler1973algebraic} --- as a proxy for network robustness.
% (Gaiarsa et al., in prep)
Algebraic connectivity relates to how difficult it is to turn a
network into completely disconnected groups of species
\citep{costa2007characterization, gibert2013spatial}. The larger the
algebraic connectivity, the more sensitive a network is to cascading
perturbations. Perturbations, such as the decrease in abundance of a
plant or pollinator, can have negative consequences for the species in
the network. For example, a decrease in abundance of a pollinator will
diminish the pollination services it provides to plants. The affected
plants would set less seeds, and decrease in abundance the subsequent
year. Consequently, other pollinators that depended on those plant
species would also be affected, and the effects of this perturbation
would continue to propagate throughout the network. Alternatively,
perturbations could also have a positive effect if, for example, the
increase in the abundance of a plant species lead to an increase in
resource availability for pollinators. The examples of negative
perturbations (e.g., resource collapse, disease spreading, parasites),
however, outnumber possible positive perturbations. It is important to
note that both robustness and algebraic connectivity assume that the
network is static. They do not account for the ability of species to
alter their interaction depending on circumstances and the resource
availability.

In order to test whether hedgerows changed in robustness as the
communities assembled, the three measures of robustness were modeled
as responses in a linear mixed model with site type as an explanatory
variable and site as a random effect \citep{lme4, lmetest}.

\section*{Results}
\label{sec:results}

Over eight years and $747$ samples, we collected and identified
$19,547$ wild bees and syrphids comprising $173$ species from $50$
genera. We observed $1,521$ unique interactions between plants and
pollinators.

\subsection*{Change point analysis}
\subsubsection*{Identifying change points}

The majority ($76\%$) of the sites underwent at least one significant
interaction reorganization (Fig.~\ref{fig:changePoints},
\ref{fig:changePoints2}).  All five of the assembling hedgerows
experienced network changing points, whereas only $40\%$ and $81\%$ of
non-assembling hedgerows and field margins, respectively, underwent
significant interaction reorganizations. Assembling hedgerows had
significantly more changing points than the non-assembling networks
(estimate of the difference in the odds ratios between assembling and
non-assembling networks, $3.316$, $95\%$ CI $[1.314, 8.572]$,
$p$-value= $0.0117$). Network assembly following restoration is thus
punctuated by more interaction reorganizations than would be expected
by external factors such as environmental shifts that would have
affected all networks (assembling or non-assembling) similarly.

\subsubsection*{Characteristics of species that contribute to change
  points}

In contradiction to the predictions of assembly by preferential
attachment, both pollinator persistence and degree were positively
related to network position variability (Fig.~\ref{fig:cv}, estimate
of the slope of closeness centrality variability and persistence $\pm$
standard error of the estimate, $0.496 \pm 0.182$, $p$-value=$0.016$;
slope of closeness centrality variability and degree, $0.004 \pm
0.002$, $p$-value=$0.056$). The slope of these relationships remained
significant when the species with the top $10$ persistence and degree
scores were dropped. In addition, plant persistence and degree were
not significantly related to network position variability
(Fig.~\ref{fig:cv}). % estimate of the slope of closeness variability and
% persistence $\pm$ SE, $-2.063 \pm 3.091$, $p$-value=$0.5$; slope of
% closeness variability and degree, $0.0018 \pm 0.002$,
% $p$-value=$0.3$).
The variability of species network position was not the result of
closeness linearly increasing through time, and, in fact, plant and
pollinator closeness decreased slightly through time (Fig.~S2,
estimate of the slope of closeness through time $\pm$ SE, pollinators:
$-0.082 \pm 0.021$, $p$-value=$0.003$; plants $-0.089 \pm 0.044$,
$p$-value=$0.04$).

\subsection*{Species and interaction turnover}
The rates of plant, pollinator and interaction temporal turnover were
similar across assembling hedgerows, non-assembling hedgerows and
field margins, though mature hedgerows had marginally less pollinator
turnover than field margins (Fig.~\ref{fig:beta}, estimate $\pm$ SE of
the difference in turnover between field margins and mature hedgerows,
$-0.051 \pm 0.026$, $p$-value=$0.049$). The turnover of plants and
interactions was not significantly different between site types. When
interactions were weighted by their similarity, both assembling and
mature hedgerows had higher rates of turnover than field margins
(Fig.~\ref{fig:beta}, estimate $\pm$ SE of the difference in turnover
between field margins and assembling hedgerows, $0.115 \pm 0.027$,
$p$-value=$0.0002$; field margins and mature hedgerows, $0.082 \pm
0.024$, $p$-value=$0.002$). The weighted interaction turnover at
assembling hedgerows, however, was not significantly higher than in
non-assembling, mature hedgerows.

\section*{Temporal changes in interaction patterns}
\subsection*{Network structure}
%Network nestedness significantly increased with assembly
%(Fig.~\ref{fig:baci}, estimate of the slope of nestedness through time
%$\pm$ SE, $1.834 \pm 0.6142$, $p$-value=$0.022$). 
Nestedness and modularity did not change in any predictable pattern
with assembly (Fig.~\ref{fig:baci}). All of the networks were
significantly nested ($z$-scores $>$ $1.96$), but not modular ($z$-scores
$<$ 2, Fig.~\ref{fig:baci}).

%though the slope was not significantly different from
%zero.(estimate
% of the slope of modularity through time $\pm$ standard error of the
% estimate, $-0.524$ $\pm$ $0.295$, $p$-value=$0.124$).
%In addition, none of the networks were significantly modular
% Fig.~\ref{fig:baci}).

Connectance decreased as the community assembled (Fig.~\ref{fig:baci},
estimate of the slope of connectance through time $\pm$ standard error
of the estimate, $-0.023$ $\pm$ $0.008$,
$p$-value=$0.007$). Specialization also decreased, though the slope
was not significantly different from
zero. %(estimate of the slope of specialization through time $\pm$ SE, $-0.926$ $\pm$ $0.450$, $p$-value=$0.078$).
Most communities were more generalized than expected when interactions
were randomized (Fig.~\ref{fig:baci}).

Both plant and pollinator species richness increased through time
(Fig.~\ref{fig:baci}, estimate of the slope of richness through time
$\pm$ SE, pollinators: $0.136$ $\pm$ $0.064$, $p$-value=$0.034$;
plants: $0.140$ $\pm$ $0.060$, $p$-value=$0.012$). Unsurprisingly,
pollinator species are colonizing and persisting at the assembling
hedgerows. Plant species richness in the networks is based on the
flowers actually visited by pollinators and not the presence of a
particular plant species at a site. Thus, though some new plant
species may establish themselves in the hedgerows, the increase in
plant richness in the networks is likely due to previously unvisited
plants attracting visitors as they mature and offer better rewards.

\subsection*{Network robustness}
Assembly did not affect network robustness to species
extinction when species were removed incrementally by
degree or abundance % (estimate
% of the slope of robustness through time $\pm$ SE, $6*10^{-5} \pm
% 4*10^{-3}$, $p$-value=$0.987$) ($0.001 \pm 0.003$,
% $p$-value=$0.65$)
In contrast, the sensitivity of networks to cascading perturbations,
as measured by the algebraic connectivity of the network, increased as
the network assembled (Fig.~S3), estimate of the slope of
sensitivity to cascading perturbations through time $\pm$ SE, $0.6814
\pm 0.272$, $p$-value=$0.042$).

\section*{Discussion}
\label{sec:discussion}

We show that the temporal assembly of plant-pollinator networks
following restoration is a highly dynamic process in which
interactions often undergo significant reorganizations, or changing
points. If these network reorganizations were a product of
environmental forces alone, we would expect to observe the same
changing points at the same time periods, consistently across all
sites. However, network changing points in non-assembling communities
are less frequent, and there are few consistent trends in the years
when change points occurred across all sites. Several sites had
network changing points between years 2009 and 2011
(Fig.~\ref{fig:changePoints}). In California, 2011 marked the
beginning of a multi-year drought. The assembling hedgerows were not
sampled in 2010, so we cannot disentangle whether the point changes
are due to skipping a year of monitoring or the
drought. Interestingly, most assembling hedgerows did not undergo a
significant interaction reorganization immediately after planting
(i.e., the transition from weedy field margin to hedgerow). This
result is consistent with the finding that hedgerow restoration takes
several years to have an impact on the plant-pollinator communities in
our study system, as well as with the observation that hedgerows do
not begin to produce many flowers until $3-5$ years following planting
\citep{kremen-2015-602}.

In a preferential attachment process, we expect that the most
persistent and high degree species (the species with the most unique
interaction partners) would remain stable in the network core during
assembly \citep{barabasi1999emergence}. Surprisingly, however, we
encountered the opposite pattern. For example, the four most
ubiquitous species in our study landscape --- \textit{Halictus
  ligatus}, \textit{Halictus tripartitus}, \textit{Lasioglossum
  (Dialictus) incompletum}, and \textit{Toxomerus marginatus} --- were
the only species that changed which module they were a member in
across years in all the assembling hedgerows. Becaused species degree
and persistence were strongly correlated, it is difficult to
disentangle the causal mechanism for why species with those
characteristics are so variable in their network position. Species
that can interaction with a relativly high number of interaction
partners may be better able to exploit the limited floral resources in
the intensively managed agriculture landscape, and thus also be the
most persistent \citep[in ant-plant mutualisms,
][]{diaz2010changes}. More persistent species usually have longer
phenologies, so they can visit many different flowers, resulting in a
higher degree \citep{Vazquez2009, fort2016abundance}. Either way, our
result suggests that adaptable species can change their network
position to utilize the most advantageous floral resources available,
which may depend on both the other pollinator species that are present
and the state of the plant community \citep{macleod2016measuring,
  gomez2006ecological, Waser1996}. Thus given the opportunity and
ability to use different resources, species will often change their
network positions \citep{macleod2016measuring}.

Interestingly, though assembling hedgerows had more network
reorganizations than non-assembling communities, pollinator species
and interaction turnover occurred at similar rates across site types.
Assembling hedgerows have higher turnover than non-assembling field
margins only when interactions were weighted by their similarity. This
is likely because though species and interactions are turning over at
the field margins, species and interactions that fill similar roles in
the network are replacing each other. In contrast, at the assembling
hedgerows, unique interactions are turning over as the networks
continually reorganize. Non-assembling mature hedgerow communities,
however, had similar rates of weighted interaction turnover as
assembling hedgerows but also the lowest pollinator
turnover. Pollinator communities at mature hedgerows may generally be
more stable, but rare and/or specialized pollinators could generate
this pattern if they entered a community, formed unique interactions
with plants that did not previously share pollinators, but did not
persist in the networks. These species would not contribute strongly
to network reorganization or species turnover, but would enhance
weighted interaction turnover. Mature hedgerows therefore both support
more stable pollinator communities, while also providing resources for
rare and/or specialized species \citep{kremen-2015-602,
  mgonigle-2015-x}.

When we explored how network-level interaction patterns changed
through time, we found few patterns. The connectance decreased, as
would be expected if the network is being colonized by specialist
species \citep{mgonigle-2015-x}. Network specialization, however, did
not change predictably with assembly. Decreasing connectance without
an accompanying increase in specialization would be possible if the
increased colonization of specialized species was accompanied by an
increase in the diet breath of resident species. This would be
expected if resident species were able to minimize their foraging time
by expanding their diet breath as plant diversity increases with
hedgerow maturation \citep{Waser1996, pyke1984optimal, Bluthgen2007,
  albrecht2010plant}.

Interestingly, however, the changes in network patterns associated
with assembly did not affect the hedgerow network robustness to
species extinction or susceptibility to cascading perturbations. The
hedgerows were designed to provide floral resources for the largest
number of pollinators across the growing season
\citep{menz-2010-4}. The generalized nature of the floral community
may explain why the networks tended to be more generalized than
expected if interactions were randomly distributed across species
(Fig.~\ref{fig:baci}). In addition, the design of the hedgerow
plantings may have facilitated the emergence of a single, highly
connected module in all of the networks (see \ref{fig:changePoints2}
for examples). This network configuration results in short path
lengths (the distance between species in a network based on their
shared partners), and thus, a perturbation in one species can more
easily spread to other species. These networks are also vulnerable to
the extinction of highly connected species \citep{Burgos2007}. In
order to promote more resilient communities, future restoration
efforts should explore designing floral communities to promote more
interaction partitioning using, for example, algorithms to optimize
different network properties based on prior knowledge of pollinator
floral preferences \citep{mgonigle2016tool}, and on desired network
architectures that renders them more robust both to species loss and
to cascading effects.

% The relationship between diversity and stability in networks
% has been the subject of considerable debate
% \citep[e.g.,][]{may1972will, pimm1984complexity,
%   montoya2006ecological}. Our results provide one of the few empirical
% examples of how restoring species diversity contributes to enhancing
% network stability.

% providing further evidence that hedgerows are a valuable tool for
% promoting species conservation and ecosystem provision in agricultural
% areas \citep{mgonigle-2015-x, ponisio2015farm,
%   kremen-2015-602}.

In general, plant-pollinator networks are highly dynamic, with high
turnover of species and interactions both within and between seasons
\citep{Burkle2011}. Though our non-assembling communities experienced
fewer network reorganizations than the assembling hedgerows, $82\%$ of
field margins and $40\%$ of mature hedgerows underwent at least one
changing point in network structure. Pollinators are also highly
opportunistic \citep{petanidou-2008-564, Vazquez2005b,
  albrecht2010plant}, though trait complementarity such as tongue
length and corolla depth impose some biophysical limits to the
interactions between plants and pollinators
\citep{Vazquez2009evaluating, Vazquez2009, Stang2009, Stang2006,
  Santamaria2007}. Such opportunism may buffer plant-pollinator
communities from global change \citep[e.g.,][]{ramos2012topological,
  kaiser2010robustness}, but our limited understanding of the assembly
of these communities impedes making such predictions
\citep{Vazquez2009, Burkle2011}. Unlike in the broader food web
literature, we have few assembly models of mutualistic network
assembly \citep{valdovinos2013adaptive, Nuismer2013, Guimaraes2011}.
In addition, the few developed models often borrow their mechanisms
from competitive interactions, leading to inaccurate biological
assumptions \citep{holland2006comment}. We need further development of
mechanistic models of mutualistic systems to generate testable
predictions, along with empirical exploration of network
assembly. Plant-pollinator communities and mutualisms are vital for
biodiversity maintenance and essential ecosystem service provision. We
must therefore understand the processes underlying their assembly to
facilitate restoration and conservation.

\section*{Acknowledgments}
\label{sec:acknowledge}

We would like to thank Paulo R.~Guimar{\~a}es Jr., Aaron Clauset,
Matthew Hutchinson, Rebecca Brunner, and Ana Paula A. Assis for their
invaluable discussions and comments, and Leto Peel for help with the
change point analysis.  We thank the growers and land owners that
allowed us to work on their property. We also greatly appreciate the
identification assistance of expert taxonomists Martin Hauser, Robbin
Thorp and Jason Gibbs.  This work was supported by funding from the
Army Research Office (W911NF-11-1-0361 to CK), the Natural Resources
Conservation Service (CIG-69-3A75-12-253, CIG-69-3A75-9-142,
CIG-68-9104-6-101 and WLF-69-7482-6-277 to The Xerces Society), the
National Science Foundation (DEB-0919128 to CK), The U.S.  Department
of Agriculture (USDA-NIFA 2012-51181-20105 to Michigan State
University).  Funding for LCP was provided by an NSF Graduate Research
Fellowship, the USDA NIFA Graduate Fellowship, and the Berkeley
Institute for Data Science. Funding for MPG was provided by S{\~a}o
Paulo Research Foundation (FAPESP, grant 2013/13319-5). We also
appreciate the Santa Fe Institute for facilitating this international
collaboration.

\bibliographystyle{ecol_let}

\bibliography{refs}

\clearpage
\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{figures/scheme.pdf}
  \caption{Diagram illustrating the analysis to examine the temporal
    turnover of interactions weighted based on their similarity. A, B
    and C are animal species, and Ps are plant species. The dendrogram
    depicts the interaction similarity across years based on the
    number of shared constituent species.}
  \label{fig:methods}
\end{figure}
\clearpage

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{../analysis/changePoint/plotting/networksv1.pdf}
  \caption{Assembling hedgerow networks had more changing points
    (vertical red lines) than non-assembling hedgerows and weedy field
    margins (a representative sample of non-assembling sites are
    depicted here). In each network, plants and pollinators are
    represented by green and yellow circles, respectively, weighted by
    their degree. Each species has a consistent position in the
    perimeter of the network across years. Asterisks indicate the year
    the hegdgerow was planted. Before that, the sites were weedy field
    margins.}
  \label{fig:changePoints}
\end{figure}
\clearpage

\begin{sidewaysfigure}
  \centering
  \includegraphics[width=.95\textwidth]{../analysis/changePoint/plotting/communitiesSkanky3.pdf}
  \caption{The species module membership between network changing
    points. Two representative assembling hedgerows are depicted. In
    the top panel, species are grouped by module. The bottom panels
    visualize the flow of species between modules between changing
    points. Each line represents the change of s species from one
    module to another. }
  \label{fig:changePoints2}
\end{sidewaysfigure}
\clearpage

\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{../analysis/variability/figures/cv/occ_degree.pdf}
  \caption{The variation coefficient of network position, as
    represented by closeness, plotted against pollinator persistence
    and degree. Persistence and degree were positively related to
    network position variability in pollinators, but unrelated in
    plants. Points represent means for each species across sites. The
    solid line indicates the mean slope estimate and the dashed lines
    are the $95\%$ confidence intervals around the estimate. }
  \label{fig:cv}
\end{figure}
\clearpage

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{../analysis/variability/figures/turnover_panels.pdf}
  \caption{Species interaction and weighted interactions turnover
    of plant-pollinator networks at non-assembling field margins
    sites, assembling hedgerows, and non-assembling, mature
    hedgerows. Rates of species and interaction turnover were similar
    between site types, though mature hedgerows has marginally
    significantly less pollinator turnover.  However, when
    interactions were weighted by their similarity, both hedgerow
    types had higher turnover than unrestored field margins. Boxplots
    represent medians (black horizontal line) first and third
    quartiles (box perimeter) and extremes (whiskers). }
  \label{fig:beta}
\end{figure}
\clearpage

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{../analysis/networkLevel/figures/baci.pdf}
  \caption{Nestedness, plant richness and pollinator richness
    increased as the networks assembled. Specialization and modularity
    remained consistent across years, while connectance decreased. The
    nestedness, modularity and specialization scores represent
    $z$-scores. Scores greater than $\sim 2$ or less than $\sim -2$
    are significantly more or less structured than randomly assembled
    networks. Points are the metric value for each site at each year
    of assembly. The solid line indicates the mean slope estimate and
    the dashed lines are the $95\%$ confidence intervals around the
    estimate.}
  \label{fig:baci}
\end{figure}
\clearpage

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End: