### 2016-12-02 ms/hardcoded/ms.tex

\documentclass[12pt]{article}
\usepackage[pdftex]{graphicx}
\usepackage[numbers]{natbib}
% \usepackage{natbib}
\usepackage{color}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{mathpazo}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{lscape}
\usepackage{array}
\usepackage{paralist}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{times}
\usepackage{textcomp}
\usepackage{xr}
\usepackage{etoolbox}
\usepackage{filecontents}
\usepackage{url}
\usepackage{rotating}
\externaldocument{SI}

\makeatletter
\def\@biblabel#1{#1.\\}
\makeatother

\parindent=2em
\setlength{\parskip}{1ex}

\RequirePackage{lineno}

\def\title{Major network reorganizations punctuate the assembly of
  plant-pollinator communities}

\def\author{Lauren C.\
  Ponisio$^{1,2,3}$, Marilia P. Gaiarsa$^4$, Claire Kremen$^1$}

\def\runninghead{Plant-pollinator network assembly}
\def\keywords{community assembly, change points, robustness,
  nestedness, modularity, restoration, mutualism, preferential attachment}

\def\extras{
  \begin{itemize}
  \item Submitted as a Letter
  \item Abstract word count: 136
  \item Main text word count: 4680
  \item Number of references: 67
  \item Number of figures: 6
  \item Author contributions: CK designed the study, LCP and CK
    collected data, LCP and MPG analyzed output data.  LCP wrote the
    first draft of the manuscript, and all authors contributed to
    revisions.
  \item Data accessibility statement: Data will de depositied in Dryad
    upon publication
  \end{itemize}
}

\def\affiliation{
  \begin{enumerate}
  \item Department of Environmental Science, Policy, and Management\\
    University of California, Berkeley\\
    130 Mulford Hall\\
    Berkeley, California, USA\\
    94720\\
  \item Berkeley Institute for Data Science (BIDS)\\
    University of California, Berkeley\\
    190 Doe Library \\
    Berkeley, California, USA\\
  \item Department of Entomology\\
    University of California, Riverside\\
    417 Entomology Bldg.\\
    Riverside, California, USA\\
    92521\\
  \item Departamento de Ecologia\\
    Universidade de S{\~a}o Paulo\\
    S{\~a}o Paulo, SP, Brazil\\
    05508-900\\
  \end{enumerate}
}

\newcommand{\mstitlepage}{
  \paragraph{Running head:} \textsc{\runninghead}
  % \parindent=0pt
  \begin{center}%
    {\LARGE \title \par}%
    \vskip 3em%
    {\large
      \lineskip .75em%
      \begin{tabular}[t]{c}%
        \author
      \end{tabular}\par}%
    \vskip 1.5em%
  \end{center}\par
  \affiliation
}
\clearpage

\begin{document}

\mstitlepage
\doublespacing
\linenumbers
\clearpage

\begin{abstract}  
  Species and interactions are being lost at alarming rates and it is
  imperative to understand how communities assemble if we are to
  prevent their collapse and restore lost interactions. Using an
  eight-year dataset comprising nearly $20,000$ pollinator visitation
  records, we explore the assembly of plant-pollinator communities at
  native plant restorations sites in an agricultural landscape. We
  find that species occupy highly dynamic network positions through
  time, causing the assembly process to be punctuated by major network
  reorganizations. The most persistent and generalized pollinator
  species are also the most variable in their network positions,
  contrary to what preferential attachment --- the most widely studied
  theory of network assembly --- predicts. Our results fundamentally
  alter our understanding of how communities assembly and how species
  interactions change through time while helping to inform efforts to
  re-assemble robust communities through restoration.
\end{abstract}

\keywords

\clearpage

\section*{Introduction}
\label{sec:introduction}

Species in ecological communities are linked through ecological
interactions. As the world continues to lose species at an alarming
rate \citep{dunn2009sixth, barnosky2011has}, anticipating a
community's ability to resist collapse will depend on the strength of
each interaction network and the species within it \citep{Memmott2004,
  Rezende2007, Bascompte2009, Thebault2010}. Furthermore, it has been
show that biodiversity has a pivotal role in maintaining ecosystems
services \citep{cardinale-2012-59}. Thus, to safeguard ecological
function, it has become increasingly imperative to aid the recovery of
lost interactions and component biodiversity and by facilitating the
assembly of robust interaction networks via ecological restoration
\citep{menz-2010-4}. However, we know very little about how ecological
networks assemble in general, rendering restoration of interacting
communities quite difficult.

Preferential attachment, the most widely explored mechanism of network
assembly, predicts that species entering a network are more likely to
interact with species that are already well-connected \citep[the
``rich-get-richer'' principle,][]{barabasi1999emergence}. In
pollination systems --- a particularly ubiquitous mutualism
\citep{ollerton-2011-321, klein-2007-303} --- some studies have found
support for this assembly mechanism. Investigating the day-to-day,
temporal assembly of a plant-pollinator network within a season,
\cite{Olesen2008} found that phenologically new plant and pollinator
species tended to interact with already well-connected species,
potentially because these species are either more abundant or
temporally persistent. Using a space-for-time substitution to study
primary succession, \cite{albrecht2010plant} also found evidence that
assembly along a glacier foreland occurred through preferential
attachment. Specifically, network nestedness (i.e, a core group of
generalists interacts with both specialist and generalist species)
increased as the community aged \citep{albrecht2010plant}. An increase
in nestedness could have occurred via the preferential attachment
process, whereby specialist species attach to the well-connected,
generalist core.

Significant reorganizations of interactions --- change points --- can
also punctuate assembly \citep{peel2014detecting}. A change point is
caused by a merge, split, fragmentation or formation of modules within
a network. Such interaction reorganizations are observed in social
networks responding to abrupt shifts in the behavior of interactors
\citep{peel2014detecting}. In ecological communities, such shifts may
occur if, as new species colonize, resident species change their
interaction partners to optimize their foraging effort
\citep{pyke1984optimal, Bluthgen2007, valdovinos2010consequences,
  albrecht2010plant, valdovinos2013adaptive}. For preferential
attachment to be at play, change points would need to be driven
primarily by peripheral, temporally variable species while a stable,
well-connected core of species remained stable in their
interactions. No studies, however, have examined if network changing
points occur during ecological network assembly and/or how these
changes relate to species behavior.

Understanding network assembly is particularly relevant to ecological
restoration, which is essentially ``applied succession''
\citep[e.g.,][]{parker1997scale}.  In pollination systems, time has
been shown to affect the structure of networks in restored areas
\citep{forup-2008-742, devoto2012understanding}, suggesting that
interactions change as a community develops. Facilitating network
restoration is especially imperative in areas where species
interactions provide essential ecosystem services, such as crop
pollination in agricultural landscapes. To promote pollinator services
in agriculture, some farmers plant strips of native plants along farm
edges (hedgerows).  By providing habitat, hedgerows augment the
richness, abundance and trait diversity of pollinators in agricultural
landscapes \citep{morandin-2013-829, kremen-2015-602,
  ponisio2015farm}, and promote the persistence and colonization of
floral resource specialists \citep{mgonigle-2015-x}. As the community
assembles, it is important to understand how these new species are
incorporated into the network as well as the consequences of adding
species for interaction patterns and robustness.

We explore the process of network development using a eight-year
dataset of plant-pollinator community assembly following hedgerow
installment in the highly simplified and intensively managed
agricultural landscape of California's Central Valley. We first
determine whether network assembly was punctuated by significant
reorganizations of interactions. We next tested whether the species
that are most variable in their network position --- and thus
important contributors to network reorganizations --- are less
persistent and connected species, as would be expected by preferential
attachment. To further explore the mechanisms underlying the temporal
dynamics of the networks, also we examine patterns in the species and
interaction temporal turnover. Finally, we investigate whether
networks assemble toward predictable interaction patterns, and the
ramifications for the robustness of the networks.

\section*{Materials \& Methods}
\label{sec:methods}

\subsection*{Study sites and collection methods}
\label{sec:study-sites}

Study sites were located in the Central Valley of California in Yolo,
Colusa and Solano Counties. This area is composed of intensively
managed agriculture --- primarily monocultures of conventional row
crops, vineyards and orchards. Hedgerows border large
(ca. $30$-hectare) crop fields and measured between $3-6$ m wide and
approximately $350$ m long. Hedgerows consist of native, perennial,
shrub and tree plantings \citep[Fig.~S1, ][]{menz-2010-4,
  kremen-2015-602, mgonigle-2015-x}.

We selected five farm edges to be restored as hedgerows. We monitored
those locations before the hedgerow was planted, and tracked the
assembly of the community for up to seven years as the hedgerow
matured (from $2006-2014$). For controls, we concurrently monitored
non-assembling communities within mature hedgerows (greater than $10$
years since planting, $N=29$) and unrestored, weedy field margins
($N=19$). The mean distance between monitoring sites was $15$ km, and
the minimum distance between sites sampled in the same year was $1$
km. The entire area surveyed spanned almost $300$ $km^2$. The crop
fields adjacent to all sites were similarly managed as intensive,
high-input monoculture.

Sites were sampled between two and five times per year (Tables S1-S3,
mean $3.4$ samples per year). In each round of sampling, the order in
which sites were sampled was randomized. Surveys were conducted under
sunny conditions when the temperature was above $21^{\circ}\mathrm{C}$
and wind speed was below $2.5$ meters/second.

During each sampling round, flower-visitors to plants in assembling
hedgerows and non-assembling controls were netted for one hour of
active search time (the timer was paused when handling specimens). The
identify of the plant being visited was recorded for each floral
visitor. Honeybees (\textit{Apis mellifera}) were not collected
because their abundance is determined largely by hive placement by
bee-keepers. All other insect flower visitors that touched the
reproductive parts of the flower were collected; however, here we
focus only on wild bees and syrphids --- the most abundant and
effective pollinators in the system (representing 49 and 19 percent of
records, respectively, C.~Kremen, A.~Klein and L.~Morandin,
unpublished data). Bee and syrphid specimens were identified to
species (or morpho-species for some bee specimens in the genera
\textit{Nomada} and \textit{Sphecodes}) by expert taxonomists.

Quantitative networks were generated for each site through
time. Because the number of sampling rounds varied between years
(Tables S1-S3), we used the mean of the interactions between a pair of
plants and pollinators within a year to represent interaction
frequency.

\subsection*{Change point analysis}
\subsubsection*{Identifying change points}
We employed a change point detection method \citep{peel2014detecting}
to identify fundamental reorganizations in large-scale interaction
patterns. Change point detection methods have yet to be generalized to
quantitative networks, so for this analysis we focused on qualitative
(binary) networks. Following \cite{peel2014detecting}, we first
defined a probability distribution over the networks using the
generalized hierarchical random graph model (GHRG). The GHRG model
captures both assortative and disassortative structure patterns at all
scales in the network \citep{peel2014detecting}. A network $G$ is
composed of vertices $V$ and edges $E$. The GHRG model decomposes the
$N$ vertices into a series of nested groups, the relationships among
which are represented by the dendrogram $T$. The tips of $T$ are the
vertices of $G$, and the probability that two vertices $u$ and $v$
connect is given by the parameter $p_r$. The probability distribution
of the network $G$ is thus defined as:

\begin{equation}
  \label{eq:lik}
  P(G|T,{pr}) = p_r^{E_r}(1-p_r)^{N_r-E_r}
\end{equation}
% 
Where $E_r$ is the observed number of edges between vertices with the
common ancestor $r$, and $N_r$ is the total possible edges, and the
rest of the notation is described above.

Using Bayesian posterior inference and techniques from phylogenetic
tree reconstruction, we fit the GHRG model to the networks
\citep{peel2014detecting}. We used a Markov chain Monte Carlo (MCMC)
procedure to first sample the posterior distribution of bipartitions,
from which a consensus tree was derived \citep{peel2014detecting}. We
used $\beta$ distributions with the hyperparameters $\alpha=\beta=1$
to define priors for $p_r$.

Once the GHRG model was fit to the networks, we determined whether a
change point occurred between two time slices. To detect a change
point, we used Bayes factors to compare the fit of two models --- one
where a change point occurred between two networks, and one where no
change occurred. We chose a sliding window of length, $w$, of four
years to detect change points. Larger windows allow for more gradual
changes, and four was the maximum possible with our eight years of
data. Lastly, to calculate a $p$-value for the Bayes factors, we used
parametric bootstrapping to numerically estimate the null distribution
\citep{peel2014detecting}. We employed code published online by
L.~Peel for the change point analysis. Analyses were conducted in
Python 3.4.

We next tested whether the change points occurring in maturing
hedgerows were a component of the assembly process or a product of
environmental shifts that lead to network reorganizations in all types
of communities. We modeled the number of change points as successes
and the total number of sample years at each site as trials, and used
a generalized linear model with Binomial error to test whether the
probability of a change point occurrence varied by site type. We used
standard techniques to determine whether the assumptions of the models
were met for this and all subsequent models. For the non-assembling
controls, only sites with five or greater survey years were included
in this analysis ($N=11$). All statistical analyses were conducted in
R 3.2.3 \citep{R}.

\subsubsection*{Characteristics of species that contribute to change
  points}

To further elucidate the nature of the change points, we examined the
characteristics of the species that contributed to interaction
reorganization. We tested whether the more persistent species with the
highest degree (number of different interaction partners) were the
most stable in their network positions, as would be expected if the
networks were assembling via preferential attachment.

We calculated species persistence as the proportion of surveys in
which a plant or pollinator is observed. Species observed consistently
within and between years were thus maximally persistent. Species
degree was calculated from interaction observations from an extensive
dataset from Yolo County (approx.~18000 interaction records) that
included both the data included in this study and additional data from
sites where we collected flower visitors using the same methods
\citep{mgonigle-2015-x, ponisio2015farm}. To represent network
position variability, we computed the coefficient of variation of
weighted closeness centrality \citep{freeman1978centrality} at each
site through time. Closeness centrality represents the importance of a
species by calculating the path lengths to other vertices (species) in
the network \citep{freeman1978centrality}. The shorter the mean path
length to other species, the higher the closeness centrality. We used
linear mixed models to test whether the species closeness variability
(log) is related to the persistence or degree of that species
\citep{lme4, lmetest}. We included random effects for species and
site. Because the degree and persistence of pollinators were strongly
correlated, ($\rho = 0.071$, $p$-value $<$ $2*10^{-16}$), we included
each explanatory variable in separate models. Plant degree and
persistence were not significantly correlated, but we used the same
models as we did for the pollinators for consistency.  Because an
approximately logarithmic increase in closeness centrality --- as
would be expected with assembly by preferential attachment --- we also
tested whether log closeness centrality increased through time.

\subsection*{Species and interaction turnover}

Reorganizations of network structure can be the result of species
turnover or species changing their interaction partners (i.e.,
re-wiring). To better understand the mechanisms underlying the
temporal dynamics of the assembling networks, we examined patterns of
species and interaction turnover. For example, assembling networks may
have higher rates of pollinator turnover than non-assembling
communities because new pollinator species are colonizing and
establishing themselves \citep{mgonigle-2015-x}. Similarly, because
species are turning over and pollinators are trying to maximize their
foraging efficiency based on the species present, interactions may
turnover more quickly than in established communities. In addition, at
assembling hedgerows, plants that are unvisited in early years may
appear to ``colonize'' the networks as they became more attractive
resources and establish new interactions with pollinators.

To estimate the temporal species and interaction turnover, we used an
approach similar to calculating spatial $\beta$-diversity. Instead of
calculating the variation in community composition across sites within
a year, we estimated turnover across years at a site. We first
calculated the pairwise dissimilarity of plants, pollinators and
interactions between years within each site using the Chao
dissimilarity estimator that incorporates abundances, while also
accounting for unobserved records \citep{chao-2005-148}. Dissimilarity
estimates can be affected by the total number of species and
individuals sampled at a site \citep[e.g.,][]{ponisio2015farm}. For
example, the probability that two sites do not share any species is
higher when there are few individuals at those sites. Following
\cite{ponisio2015farm}, we used null models that constrained species
richness to estimate the deviation of the observed dissimilarity from
that which would be expected under a random community assembly
process. With the corrected dissimilarity values, we then calculated
the multivariate dispersion of community composition across years
\citep{anderson-2011-19}. In order to test whether assembling
hedgerows had more species and interactions turnover than
non-assembling communities, the species and interaction temporal
turnover estimates were modeled as responses in a linear mixed model
with site type as an explanatory variable and site as a random effect
\citep{lme4, lmetest}.

Though species may turnover across years, some groups of species may
essentially replace each other if they fill similar roles in the
network, occupying the same network position and interacting with
similar species. At non-assembling communities, species turnover may
overestimate the temporal changes in the networks if the interactions
occurring in one year are similar to those in the next year when they
are weighted by the similarity of their constituent species
(Fig.~\ref{fig:methods}). We developed a method to examine the
temporal turnover of interactions with weightings based on their
similarity. We followed the algorithm of \cite{ahn2010link} to cluster
all the interactions (edges) hierarchically across sites and years
based on their similarity, and built a dendrogram. The interaction
similarity is based how many plants and pollinators (vertices) two
edges share \citep{ahn2010link, kalinka2011linkcomm}. The more species
edges shared in common, the shorter the branch length between them on
the dendrogram.  We next calculated the temporal turnover of
interactions weighted by their similarity, as approximated by
``phylogenetic'' distance \citep{graham2008phylogenetic,
  picante-2010-1463}. We then used linear mixed models to test whether
the weighted turnover of interactions varied between assembling and
non-assembling networks, including site as a random effect
\citep{lme4, lmetest}.

\section*{Temporal changes in interaction patterns}

\subsection*{Network structure}
Any changing points in network structure may contribute to the
reorganization of the assembling networks into predictable interaction
patterns. Pollination networks exhibit two main structural patterns
--- modularity \citep[e.g.,][]{Olesen2007} and nestedness
\citep[e.g.,][]{Bascompte2006, Bascompte2003}. In modular networks,
interactions are insular, occurring within separate groups or
``modules'' more often than between modules. Modules in the network
may fragment as the network assembles, enhancing
modularity. Conversely, nested networks are like a pyramid of
interactions, where there are some species that interact with many
species, other species that interact with a subset of those species,
and so on. If species entering the network tend to interact with the
generalist base of the network pyramid as would be expected with
preferential attachment, nestedness would increase through time. The
connectance --- the proportion of observed out of possible
interactions --- would also decrease as new, specialist species,
preferentially attach to the core. Finally, network-level
specialization will increase if specialist species colonize the
network or species begin to limit their interaction niche breath as
the network assembles \citep{bluthgen-2006-9}.

To evaluate network nestedness, we used the estimator weighted NODF
\citep{almeida-neto-2008-1227}. NODF evaluates whether species with
fewer partners interact with subsets of partners with which more
connected species interact \citep{almeida-neto-2008-1227}. To estimate
modularity, we used a hierarchical clustering algorithm
\citep{Newman2004, csardi-2006}. We evaluated network specialization
with the metric H2, which estimates the deviation of the observed
interaction frequency between plants and pollinators from a null
expectation where all partners interact in proportion to their
abundances \citep{bluthgen-2006-9}. It ranges from zero for
generalized networks to one for specialized networks.  We calculated
standardized $z$-scores so that nestedness, modularity and
specialization metrics could be compared across communities. The
$z$-scores were calculated by generating an ensemble of $999$ randomly
assembled communities, subtracting the mean of the statistic
calculated across these communities from the observed value, and then
dividing by the standard deviation. To assemble random communities, we
reshuffled the interactions between species but fixed the total number
of interactions, species and interaction frequency distributions
\citep{Galeano2009}.

To test whether network modularity, nestedness, connectance or
specialization changed linearly with assembly, we used linear mixed
models with the descriptive network metrics as the response variable,
year of assembly as the explanatory variable, and random effects of
site and year. The number of species in a network affects the patterns
of interaction possible, so we also examined the change in plant and
pollinator species richness through time. We employed generalized
linear mixed models with Poisson error to model richness
\citep{lme4}. We scaled explanatory variables.

\subsection*{Network robustness}
Finally, we tested whether the changes in interaction patterns
associated with network assembly affect the robustness of the network
to species loss and cascading perturbations. Following
\cite{Memmott2004}, we simulated plant species extinction and the
subsequent extinction cascades of pollinator species. Because the
reproduction of plant species is facilitated by active restoration
efforts, it is unlikely the extinction of pollinator species would
affect plant populations in the hedgerows. However, plants ceasing to
bloom, for example in response to drought, will likely affect the
pollinators that depend on them. We eliminated plants species based on
their degree or abundance, and then calculated the number of
pollinators that secondarily went extinct. The area below the
extinction curve is an estimate of network robustness
\citep{Memmott2004, Burgos2007, bipartite}.

We also explored how the robustness to cascading perturbations changed
as the community assembled, using algebraic connectivity --- the
second smallest eigenvalue of the Laplacian matrix
\citep{fiedler1973algebraic} --- as a proxy for network robustness.
Algebraic connectivity relates to how difficult it is to turn a
network into completely disconnected groups of species
\citep{costa2007characterization, gibert2013spatial}. The larger the
algebraic connectivity, the more sensitive a network is to cascading
perturbations. Perturbations, such as the decrease in abundance of a
plant or pollinator, can have negative consequences for the species in
the network. For example, a decrease in abundance of a pollinator will
diminish the pollination services it provides to plants. The affected
plants would set less seeds, and decrease in abundance the subsequent
year. Consequently, other pollinators that depended on those plant
species would also be affected, and the effects of this perturbation
would continue to propagate throughout the network. Alternatively,
perturbations could also have a positive effect if, for example, the
increase in the abundance of a plant species lead to an increase in
resource availability for pollinators. The examples of negative
perturbations (e.g., resource collapse, disease spreading, parasites),
however, outnumber possible positive perturbations. It is important to
note that both robustness and algebraic connectivity assume that the
network is static --- they do not account for the ability of species
to alter their interaction depending on circumstances and the resource
availability.

In order to test whether hedgerows changed in robustness as the
communities assembled, the three measures of robustness were modeled
as responses in a linear mixed model with site type as an explanatory
variable and site as a random effect \citep{lme4, lmetest}.

\section*{Results}
\label{sec:results}

Over eight years and $747$ samples, we collected and identified
$19,547$ wild bees and syrphids comprising $173$ species from $50$
genera. We observed $1,521$ unique interactions between plants and
pollinators.

\subsection*{Change point analysis}
\subsubsection*{Identifying change points}

The majority ($76\%$) of the sites underwent at least one significant
interaction reorganization (Fig.~\ref{fig:changePoints},
\ref{fig:changePoints2}).  All five of the assembling hedgerows
experienced network changing points, whereas only $40\%$ and $81\%$ of
non-assembling hedgerows and field margins, respectively, underwent
significant interaction reorganizations. Assembling hedgerows had
significantly more changing points than the non-assembling networks
(estimate of the difference in the odds ratios between assembling and
non-assembling networks, $3.316$, $95\%$ CI $[1.314, 8.572]$,
$p$-value= $0.0117$). Network assembly following restoration is thus
punctuated by more interaction reorganizations than would be expected
by external factors such as environmental shifts that would have
affected all networks (assembling or non-assembling) similarly.

\subsubsection*{Characteristics of species that contribute to change
  points}

In contradiction to the predictions of assembly by preferential
attachment, both pollinator persistence and degree were positively
related to network position variability (Fig.~\ref{fig:cv}, estimate
of the slope of closeness centrality variability and persistence $\pm$
standard error of the estimate, $0.496 \pm 0.182$, $p$-value=$0.016$;
slope of closeness centrality variability and degree, $0.004 \pm
0.002$, $p$-value=$0.056$). The slope of these relationships remained
significant when the species with the top $10$ persistence and degree
scores were dropped. In addition, plant persistence and degree were
not significantly related to network position variability
(Fig.~\ref{fig:cv}).  The variability of species network position was
not the result of closeness linearly increasing through time, and, in
fact, plant and pollinator closeness decreased slightly through time
(Fig.~S2, estimate of the slope of closeness through time $\pm$ SE,
pollinators: $-0.082 \pm 0.021$, $p$-value=$0.003$; plants $-0.089 \pm
0.044$, $p$-value=$0.04$).

\subsection*{Species and interaction turnover}
The rates of plant, pollinator and interaction temporal turnover were
similar across assembling hedgerows, non-assembling hedgerows and
field margins, though mature hedgerows had marginally less pollinator
turnover than field margins (Fig.~\ref{fig:beta}, estimate $\pm$ SE of
the difference in turnover between field margins and mature hedgerows,
$-0.051 \pm 0.026$, $p$-value=$0.049$). The turnover of plants and
interactions was not significantly different between site types. When
interactions were weighted by their similarity, both assembling and
mature hedgerows had higher rates of turnover than field margins
(Fig.~\ref{fig:beta}, estimate $\pm$ SE of the difference in turnover
between field margins and assembling hedgerows, $0.115 \pm 0.027$,
$p$-value=$0.0002$; field margins and mature hedgerows, $0.082 \pm
0.024$, $p$-value=$0.002$). The weighted interaction turnover at
assembling hedgerows, however, was not significantly higher than in
non-assembling, mature hedgerows.

\section*{Temporal changes in interaction patterns}
\subsection*{Network structure}

Nestedness, modularity and specialization did not change predictably
assembly (Fig.~\ref{fig:baci}). All of the networks were significantly
nested ($z$-scores $>$ $2$), but not modular ($z$-scores $<$ 2,
Fig.~\ref{fig:baci}). Most communities were more generalized than
expected when interactions were randomized
(Fig.~\ref{fig:baci}). Connectance decreased as the community
assembled (Fig.~\ref{fig:baci}, estimate of the slope of connectance
through time $\pm$ standard error of the estimate, $-0.023$ $\pm$
$0.008$, $p$-value=$0.007$).

Both plant and pollinator species richness increased through time
(Fig.~\ref{fig:baci}, estimate of the slope of richness through time
$\pm$ SE, pollinators: $0.136$ $\pm$ $0.064$, $p$-value=$0.034$;
plants: $0.140$ $\pm$ $0.060$, $p$-value=$0.012$). Unsurprisingly,
pollinator species are colonizing and persisting at the assembling
hedgerows. Plant species richness in the networks is based on the
flowers actually visited by pollinators and not the presence of a
particular plant species at a site. Thus, though some new plant
species may establish themselves in the hedgerows, the increase in
plant richness in the networks is likely due to previously unvisited
plants attracting visitors as they mature and offer better rewards.

\subsection*{Network robustness}
Assembly did not affect network robustness to species extinction when
species were removed incrementally by degree or abundance
(Fig.~S3). Similarly, the sensitivity of networks to cascading
perturbations, as measured by the algebraic connectivity of the
network, did not change predictably as the community assembled
(Fig.~S3).

\section*{Discussion}
\label{sec:discussion}

We show that the temporal assembly of plant-pollinator networks
following restoration is a highly dynamic process in which
interactions often undergo significant reorganizations, or changing
points. If these network reorganizations were a product of
environmental forces alone, we would expect to observe the same
changing points at the same time periods, consistently across all
sites. However, network changing points in non-assembling communities
are less frequent, and there are few consistent trends in the years
when change points occurred across all sites. Several sites had
network changing points between years 2009 and 2011
(Fig.~\ref{fig:changePoints}). In California, 2011 marked the
beginning of a multi-year drought. The assembling hedgerows were not
sampled in 2010, so we cannot disentangle whether the point changes
are due to skipping a year of monitoring or the
drought. Interestingly, most assembling hedgerows did not undergo a
significant interaction reorganization immediately after planting
(i.e., the transition from weedy field margin to hedgerow). This
result is consistent with the finding that hedgerow restoration takes
several years to have an impact on the plant-pollinator communities in
our study system, as well as with the observation that hedgerows do
not begin to produce many flowers until $3-5$ years following planting
\citep{kremen-2015-602}.

In a preferential attachment process, we expect that the most
persistent and high degree species (the species with the most unique
interaction partners) would remain stable in the network core during
assembly \citep{barabasi1999emergence}. Surprisingly, however, we
encountered the opposite pattern. For example, the four most
ubiquitous species in our study landscape --- \textit{Halictus
  ligatus}, \textit{Halictus tripartitus}, \textit{Lasioglossum
  (Dialictus) incompletum}, and \textit{Toxomerus marginatus} --- were
the only species that changed which module they were a member in
across years in all the assembling hedgerows. Because species degree
and persistence were strongly correlated, it is difficult to
disentangle the causal mechanism for why species with those
characteristics are so variable in their network position. Species
that can interact with a relatively high number of partners may be
better able to exploit the limited floral resources in the intensively
managed agriculture landscape, and thus also be the most persistent
\citep[in ant-plant mutualisms, ][]{diaz2010changes}. More persistent
species usually have longer phenologies, so they can visit many
different flowers, resulting in a higher degree \citep{Vazquez2009,
  fort2016abundance}. Either way, our result suggests that adaptable
species can change their network position to utilize the most
advantageous floral resources available, which may depend on both the
other pollinator species that are present and the state of the plant
community \citep{Waser1996, gomez2006ecological,
  macleod2016measuring}. Thus given the opportunity and ability to use
different resources, species will often change their network positions
\citep{macleod2016measuring}.

Interestingly, though assembling hedgerows had more network
reorganizations than non-assembling communities, pollinator species
and interaction turnover occurred at similar rates across site types.
Assembling hedgerows have higher turnover than non-assembling field
margins only when interactions were weighted by their similarity. This
is likely because though species and interactions are turning over at
the field margins, species and interactions that fill similar roles in
the network are replacing each other. In contrast, at the assembling
hedgerows, unique interactions are turning over as the networks
continually reorganize. Non-assembling mature hedgerow communities,
however, had similar rates of weighted interaction turnover as
assembling hedgerows but also the lowest pollinator
turnover. Pollinator communities at mature hedgerows may generally be
more stable, but rare and/or specialized pollinators could generate
this pattern if they entered a community, formed unique interactions
with plants that did not previously share pollinators, but did not
persist in the networks. These species would not contribute strongly
to network reorganization or species turnover, but would enhance
weighted interaction turnover. Mature hedgerows therefore both support
more stable pollinator communities, while also providing resources for
rare and/or specialized species \citep{kremen-2015-602,
  mgonigle-2015-x}.

When we explored how network-level interaction patterns changed
through time, we found few patterns. The connectance decreased, as
would be expected if the network is being colonized by specialist
species \citep{mgonigle-2015-x}. Network specialization, however, did
not change predictably with assembly. Decreasing connectance without
an accompanying increase in specialization would be possible if the
increased colonization of specialized species was accompanied by an
increase in the diet breath of resident species. This would be
expected if resident species were able to minimize their foraging time
by expanding their diet breath as plant diversity increases with
hedgerow maturation \citep{pyke1984optimal, Waser1996, Bluthgen2007,
  albrecht2010plant}.

Interestingly, however, the changes in network patterns associated
with assembly did not affect the hedgerow network robustness to
species extinction or susceptibility to cascading perturbations. The
hedgerows were designed to provide floral resources for the largest
number of pollinators across the growing season
\citep{menz-2010-4}. The generalized nature of the floral community
may explain why the networks tended to be more generalized than
expected if interactions were randomly distributed across species
(Fig.~\ref{fig:baci}). In addition, the design of the hedgerow
plantings may have facilitated the emergence of a single, highly
connected module in all of the networks (see \ref{fig:changePoints2}
for examples). This network configuration results in short path
lengths (the distance between species in a network based on their
shared partners), and thus, a perturbation in one species can more
easily spread to other species. These networks are also vulnerable to
the extinction of highly connected species \citep{Burgos2007}. In
order to promote more resilient communities, future restoration
efforts should explore designing floral communities to promote more
interaction partitioning using, for example, algorithms to optimize
different network properties based on prior knowledge of pollinator
floral preferences \citep{mgonigle2016tool}, and on desired network
architectures that renders them more robust both to species loss and
to cascading effects.

In general, plant-pollinator networks are highly dynamic, with high
turnover of species and interactions both within and between seasons
\citep{Burkle2011}. Though our non-assembling communities experienced
fewer network reorganizations than the assembling hedgerows, $82\%$ of
field margins and $40\%$ of mature hedgerows underwent at least one
changing point in network structure. Pollinators are highly
opportunistic \citep{petanidou-2008-564, Vazquez2005b,
  albrecht2010plant}, though trait complementarity such as tongue
length and corolla depth impose some biophysical limits to the
interactions between plants and pollinators \citep{Stang2006,
  Santamaria2007, Vazquez2009evaluating, Vazquez2009, Stang2009}. Such
opportunism may buffer plant-pollinator communities from global change
\citep[e.g.,][]{ramos2012topological, kaiser2010robustness}, but our
limited understanding of the assembly of these communities impedes
making such predictions \citep{Vazquez2009, Burkle2011}. Unlike in the
broader food web literature, we have few assembly models of
mutualistic network assembly \citep{valdovinos2013adaptive,
  Nuismer2013, Guimaraes2011}.  In addition, the few developed models
often borrow their mechanisms from competitive interactions, leading
to inaccurate biological assumptions \citep{holland2006comment}. We
need further development of mechanistic models of mutualistic systems
to generate testable predictions, along with empirical exploration of
network assembly. Plant-pollinator communities and mutualisms are
vital for biodiversity maintenance and essential ecosystem service
provision. We must therefore understand the processes underlying their
assembly to facilitate restoration and conservation.

\section*{Acknowledgments}
\label{sec:acknowledge}

We would like to thank Paulo R.~Guimar{\~a}es Jr., Aaron Clauset,
Matthew Hutchinson, Rebecca Brunner, and Ana Paula A. Assis for their
invaluable discussions and comments, and Leto Peel for help with the
change point analysis.  We thank the growers and land owners that
allowed us to work on their property. We also greatly appreciate the
identification assistance of expert taxonomists Martin Hauser, Robbin
Thorp and Jason Gibbs.  This work was supported by funding from the
Army Research Office (W911NF-11-1-0361 to CK), the Natural Resources
Conservation Service (CIG-69-3A75-12-253, CIG-69-3A75-9-142,
CIG-68-9104-6-101 and WLF-69-7482-6-277 to The Xerces Society), the
National Science Foundation (DEB-0919128 to CK), The U.S.  Department
of Agriculture (USDA-NIFA 2012-51181-20105 to Michigan State
University).  Funding for LCP was provided by an NSF Graduate Research
Fellowship, the USDA NIFA Graduate Fellowship, and the Berkeley
Institute for Data Science. Funding for MPG was provided by S{\~a}o
Paulo Research Foundation (FAPESP, grant 2013/13319-5). We also
appreciate the Santa Fe Institute for facilitating this international
collaboration.

\begin{thebibliography}{67}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{Ahn \emph{et~al.}(2010)Ahn, Bagrow \& Lehmann}]{ahn2010link}
\hspace{0em}\\Ahn, Y.Y., Bagrow, J.P. \& Lehmann, S. (2010).
\newblock Link communities reveal multiscale complexity in networks.
\newblock \emph{Nature}, 466, 761--764.

\bibitem[{Albrecht \emph{et~al.}(2010)Albrecht, Riesen \&
  Schmid}]{albrecht2010plant}
\hspace{0em}\\Albrecht, M., Riesen, M. \& Schmid, B. (2010).
\newblock Plant--pollinator network assembly along the chronosequence of a
  glacier foreland.
\newblock \emph{Oikos}, 119, 1610--1624.

\bibitem[{Almeida-neto \emph{et~al.}(2008)Almeida-neto, Guimar\~{a}es,
  Guimar{\~a}es, Loyola \& Ulrich}]{almeida-neto-2008-1227}
\hspace{0em}\\Almeida-neto, M., Guimar\~{a}es, P., Guimar{\~a}es, P., Loyola,
  R. \& Ulrich, W. (2008).
\newblock A consistent metric for nestedness analysis in ecological systems:
  reconciling concept and measurement.
\newblock \emph{Oikos}, 117, 1227--1239.

\bibitem[{Anderson \emph{et~al.}(2011)Anderson, Crist, Chase, Vellend, Inouye,
  Freestone, Sanders, Cornell, Comita, Davies \emph{et~al.}}]{anderson-2011-19}
\hspace{0em}\\Anderson, M.J., Crist, T.O., Chase, J.M., Vellend, M., Inouye,
  B.D., Freestone, A.L., Sanders, N.J., Cornell, H.V., Comita, L.S., Davies,
  K.F. \emph{et~al.} (2011).
\newblock Navigating the multiple meanings of $\beta$ diversity: a roadmap for
  the practicing ecologist.
\newblock \emph{Ecol. Lett.}, 14, 19--28.

\bibitem[{Barab{\'a}si \& Albert(1999)}]{barabasi1999emergence}
\hspace{0em}\\Barab{\'a}si, A.L. \& Albert, R. (1999).
\newblock Emergence of scaling in random networks.
\newblock \emph{Science}, 286, 509--512.

\bibitem[{Barnosky \emph{et~al.}(2011)Barnosky, Matzke, Tomiya, Wogan, Swartz,
  Quental, Marshall, McGuire, Lindsey, Maguire \emph{et~al.}}]{barnosky2011has}
\hspace{0em}\\Barnosky, A.D., Matzke, N., Tomiya, S., Wogan, G.O., Swartz, B.,
  Quental, T.B., Marshall, C., McGuire, J.L., Lindsey, E.L., Maguire, K.C.
  \emph{et~al.} (2011).
\newblock Has the earth's sixth mass extinction already arrived?
\newblock \emph{Nature}, 471, 51--57.

\bibitem[{Bascompte \emph{et~al.}(2003)Bascompte, Jordano, Meli\'an \&
  Olesen}]{Bascompte2003}
\hspace{0em}\\Bascompte, J., Jordano, P., Meli\'an, C.J. \& Olesen, J.M.
  (2003).
\newblock The nested assembly of plant-animal mutualistic networks.
\newblock \emph{Proc. Natl. Acad. Sci. USA}, 100, 9383--9387.

\bibitem[{Bascompte \emph{et~al.}(2006)Bascompte, Jordano \&
  Olesen}]{Bascompte2006}
\hspace{0em}\\Bascompte, J., Jordano, P. \& Olesen, J.M. (2006).
\newblock Asymmetric coevolutionary networks facilitate biodiversity
  maintenance.
\newblock \emph{Science}, 312, 431--433.

\bibitem[{Bascompte \& Stouffer(2009)}]{Bascompte2009}
\hspace{0em}\\Bascompte, J. \& Stouffer, D.B. (2009).
\newblock The assembly and disassembly of ecological networks.
\newblock \emph{Phil. Trans. R. Soc. B}, 364, 1781.

\bibitem[{Bates \emph{et~al.}(2014)Bates, Maechler, Bolker \& Walker}]{lme4}
\hspace{0em}\\Bates, D., Maechler, M., Bolker, B. \& Walker, S. (2014).
\newblock \emph{lme4: Linear mixed-effects models using Eigen and S4}.
\newblock R package version 1.1-7.

\bibitem[{Bl{\"u}thgen \emph{et~al.}(2006)Bl{\"u}thgen, Menzel \&
  Bl{\"u}thgen}]{bluthgen-2006-9}
\hspace{0em}\\Bl{\"u}thgen, N., Menzel, F. \& Bl{\"u}thgen, N. (2006).
\newblock Measuring specialization in species interaction networks.
\newblock \emph{BMC Ecol.}, 6, 9.

\bibitem[{Bl{\"u}thgen \emph{et~al.}(2007)Bl{\"u}thgen, Menzel, Hovestadt,
  Fiala \& Bl{\"u}thgen}]{Bluthgen2007}
\hspace{0em}\\Bl{\"u}thgen, N., Menzel, F., Hovestadt, T., Fiala, B. \&
  Bl{\"u}thgen, N. (2007).
\newblock Specialization, constraints, and conflicting interests in mutualistic
  networks.
\newblock \emph{Curr. Biol.}, 17, 341--346.

\bibitem[{Burgos \emph{et~al.}(2007)Burgos, Ceva, Perazzo, Devoto, Medan,
  Zimmermann \& Delbue}]{Burgos2007}
\hspace{0em}\\Burgos, E., Ceva, H., Perazzo, R.P., Devoto, M., Medan, D.,
  Zimmermann, M. \& Delbue, A.M. (2007).
\newblock Why nestedness in mutualistic networks?
\newblock \emph{J. Theor. Biol.}, 249, 307--313.

\bibitem[{Burkle \& Alarc{\'o}n(2011)}]{Burkle2011}
\hspace{0em}\\Burkle, L.A. \& Alarc{\'o}n, R. (2011).
\newblock The future of plant-pollinator diversity: Understanding interaction
  networks across time, space, and global change.
\newblock \emph{Am. J. Bot.}, 98, 528.

\bibitem[{Cardinale \emph{et~al.}(2012)Cardinale, Duffy, Gonzalez, Hooper,
  Perrings, Venail, Narwani, Mace, Tilman, Wardle
  \emph{et~al.}}]{cardinale-2012-59}
\hspace{0em}\\Cardinale, B.J., Duffy, J.E., Gonzalez, A., Hooper, D.U.,
  Perrings, C., Venail, P., Narwani, A., Mace, G.M., Tilman, D., Wardle, D.A.
  \emph{et~al.} (2012).
\newblock Biodiversity loss and its impact on humanity.
\newblock \emph{Nature}, 486, 59--67.

\bibitem[{Chao \emph{et~al.}(2005)Chao, Chazdon, Colwell \&
  Shen}]{chao-2005-148}
\hspace{0em}\\Chao, A., Chazdon, R.L., Colwell, R.K. \& Shen, T.J. (2005).
\newblock A new statistical approach for assessing similarity of species
  composition with incidence and abundance data.
\newblock \emph{Ecol. Lett.}, 8, 148--159.

\bibitem[{Costa \emph{et~al.}(2007)Costa, Rodrigues, Travieso \&
  Villas~Boas}]{costa2007characterization}
\hspace{0em}\\Costa, L.d.F., Rodrigues, F.A., Travieso, G. \& Villas~Boas, P.R.
  (2007).
\newblock Characterization of complex networks: A survey of measurements.
\newblock \emph{Adv. Phys.}, 56, 167--242.

\bibitem[{Csardi \& Nepusz(2006)}]{csardi-2006}
\hspace{0em}\\Csardi, G. \& Nepusz, T. (2006).
\newblock The igraph software package for complex network research.
\newblock \emph{InterJournal}, p. 1695.

\bibitem[{Devoto \emph{et~al.}(2012)Devoto, Bailey, Craze \&
  Memmott}]{devoto2012understanding}
\hspace{0em}\\Devoto, M., Bailey, S., Craze, P. \& Memmott, J. (2012).
\newblock Understanding and planning ecological restoration of
  plant--pollinator networks.
\newblock \emph{Ecol. Lett.}, 15, 319--328.

\bibitem[{D{\'\i}az-Castelazo \emph{et~al.}(2010)D{\'\i}az-Castelazo,
  Guimar{\~a}es~Jr, Jordano, Thompson, Marquis \& Rico-Gray}]{diaz2010changes}
\hspace{0em}\\D{\'\i}az-Castelazo, C., Guimar{\~a}es~Jr, P.R., Jordano, P.,
  Thompson, J.N., Marquis, R.J. \& Rico-Gray, V. (2010).
\newblock Changes of a mutualistic network over time: reanalysis over a 10-year
  period.
\newblock \emph{Ecology}, 91, 793--801.

\bibitem[{Dormann \emph{et~al.}(2008)Dormann, Gruber \& Fr{\"u}nd}]{bipartite}
\hspace{0em}\\Dormann, C., Gruber, B. \& Fr{\"u}nd, J. (2008).
\newblock Introducing the bipartite package: analysing ecological networks.
\newblock \emph{R News}, 8, 8.

\bibitem[{Dunn \emph{et~al.}(2009)Dunn, Harris, Colwell, Koh \&
  Sodhi}]{dunn2009sixth}
\hspace{0em}\\Dunn, R.R., Harris, N.C., Colwell, R.K., Koh, L.P. \& Sodhi, N.S.
  (2009).
\newblock The sixth mass coextinction: are most endangered species parasites
  and mutualists?
\newblock \emph{Proc. R. Soc. Lond. B}, 276, 3037--3045.

\bibitem[{Fiedler(1973)}]{fiedler1973algebraic}
\hspace{0em}\\Fiedler, M. (1973).
\newblock Algebraic connectivity of graphs.
\newblock \emph{Czech. Math. J.}, 23, 298--305.

\bibitem[{Fort \emph{et~al.}(2016)Fort, V{\'a}zquez \& Lan}]{fort2016abundance}
\hspace{0em}\\Fort, H., V{\'a}zquez, D.P. \& Lan, B.L. (2016).
\newblock Abundance and generalisation in mutualistic networks: solving the
  chicken-and-egg dilemma.
\newblock \emph{Ecol. Lett.}, 19, 4--11.

\bibitem[{Forup \emph{et~al.}(2008)Forup, Henson, Craze \&
  Memmott}]{forup-2008-742}
\hspace{0em}\\Forup, M., Henson, K., Craze, P. \& Memmott, J. (2008).
\newblock The restoration of ecological interactions: plant-pollinator networks
  on ancient and restored heathlands.
\newblock \emph{J. Appl. Ecol.}, 45, 742--752.

\bibitem[{Freeman(1978)}]{freeman1978centrality}
\hspace{0em}\\Freeman, L.C. (1978).
\newblock Centrality in social networks conceptual clarification.
\newblock \emph{Soc. Networks}, 1, 215--239.

\bibitem[{Galeano \emph{et~al.}(2009)Galeano, Pastor \& Iriondo}]{Galeano2009}
\hspace{0em}\\Galeano, J., Pastor, J.M. \& Iriondo, J.M. (2009).
\newblock Weighted-interaction nestedness estimator (wine): a new estimator to
  calculate over frequency matrices.
\newblock \emph{Environ. Model. Softw.}, 24, 1342--1346.

\bibitem[{Gibert \emph{et~al.}(2013)Gibert, Pires, Thompson \&
  Guimar{\~a}es~Jr}]{gibert2013spatial}
\hspace{0em}\\Gibert, J.P., Pires, M.M., Thompson, J.N. \& Guimar{\~a}es~Jr,
  P.R. (2013).
\newblock The spatial structure of antagonistic species affects coevolution in
  predictable ways.
\newblock \emph{Am. Nat.}, 182, 578--591.

\bibitem[{G{\'o}mez \& Zamora(2006)}]{gomez2006ecological}
\hspace{0em}\\G{\'o}mez, J.M. \& Zamora, R. (2006).
\newblock Ecological factors that promote the evolution of generalization in
  pollination systems.
\newblock \emph{Plant-pollinator interactions, NM Waser, J. Ollerton (eds.).
  The University of Chicago Press, Chicago}, pp. 145--166.

\bibitem[{Graham \& Fine(2008)}]{graham2008phylogenetic}
\hspace{0em}\\Graham, C.H. \& Fine, P.V. (2008).
\newblock Phylogenetic beta diversity: linking ecological and evolutionary
  processes across space in time.
\newblock \emph{Ecol. Lett.}, 11, 1265--1277.

\bibitem[{Guimar{\~a}es \emph{et~al.}(2011)Guimar{\~a}es, Jordano \&
  Thompson}]{Guimaraes2011}
\hspace{0em}\\Guimar{\~a}es, P.R., Jordano, P. \& Thompson, J.N. (2011).
\newblock Evolution and coevolution in mutualistic networks.
\newblock \emph{Ecol. Lett.}, 14, 877--885.

\bibitem[{Holland \emph{et~al.}(2006)Holland, Okuyama \&
  DeAngelis}]{holland2006comment}
\hspace{0em}\\Holland, J.N., Okuyama, T. \& DeAngelis, D.L. (2006).
\newblock Comment on ``asymmetric coevolutionary networks facilitate
  biodiversity maintenance''.
\newblock \emph{Science}, 313, 1887b--1887b.

\bibitem[{Kaiser-Bunbury \emph{et~al.}(2010)Kaiser-Bunbury, Muff, Memmott,
  M{\"u}ller \& Caflisch}]{kaiser2010robustness}
\hspace{0em}\\Kaiser-Bunbury, C.N., Muff, S., Memmott, J., M{\"u}ller, C.B. \&
  Caflisch, A. (2010).
\newblock The robustness of pollination networks to the loss of species and
  interactions: a quantitative approach incorporating pollinator behaviour.
\newblock \emph{Ecol. Lett.}, 13, 442--452.

\bibitem[{Kalinka \& Tomancak(2011)}]{kalinka2011linkcomm}
\hspace{0em}\\Kalinka, A.T. \& Tomancak, P. (2011).
\newblock linkcomm: an r package for the generation, visualization, and
  analysis of link communities in networks of arbitrary size and type.
\newblock \emph{Bioinformatics}, 27.

\bibitem[{Kembel \emph{et~al.}(2010)Kembel, Cowan, Helmus, Cornwell, Morlon,
  Ackerly, Blomberg \& Webb}]{picante-2010-1463}
\hspace{0em}\\Kembel, S., Cowan, P., Helmus, M., Cornwell, W., Morlon, H.,
  Ackerly, D., Blomberg, S. \& Webb, C. (2010).
\newblock Picante: {R} tools for integrating phylogenies and ecology.
\newblock \emph{Bioinformatics}, 26, 1463--1464.

\bibitem[{Klein \emph{et~al.}(2007)Klein, Vaissi\`{e}re, Cane,
  Steffan-Dewenter, Cunningham, Kremen \& Tscharntke}]{klein-2007-303}
\hspace{0em}\\Klein, A., Vaissi\`{e}re, B., Cane, J., Steffan-Dewenter, I.,
  Cunningham, S., Kremen, C. \& Tscharntke, T. (2007).
\newblock Importance of pollinators in changing landscapes for world crops.
\newblock \emph{Proc. R. Soc. B}, 274, 303--313.

\bibitem[{Kremen \& M'Gonigle(2015)}]{kremen-2015-602}
\hspace{0em}\\Kremen, C. \& M'Gonigle, L.K. (2015).
\newblock Small-scale restoration in intensive agricultural landscapes supports
  more specialized and less mobile pollinator species.
\newblock \emph{J. Appl. Ecol.}, 52, 602--610.

\bibitem[{Kuznetsova \emph{et~al.}(2014)Kuznetsova, {Bruun Brockhoff} \& {Haubo
  Bojesen Christensen}}]{lmetest}
\hspace{0em}\\Kuznetsova, A., {Bruun Brockhoff}, P. \& {Haubo Bojesen
  Christensen}, R. (2014).
\newblock \emph{{lmerTest}: Tests for random and fixed effects for linear mixed
  effect models (lmer objects of lme4 package).}
\newblock R package version 2.0-11.

\bibitem[{MacLeod \emph{et~al.}(2016)MacLeod, Genung, Ascher \&
  Winfree}]{macleod2016measuring}
\hspace{0em}\\MacLeod, M., Genung, M.A., Ascher, J.S. \& Winfree, R. (2016).
\newblock Measuring partner choice in plant--pollinator networks: Using null
  models to separate rewiring and fidelity from chance.
\newblock \emph{Ecology}.

\bibitem[{Memmott \emph{et~al.}(2004)Memmott, Waser \& Price}]{Memmott2004}
\hspace{0em}\\Memmott, J., Waser, N.M. \& Price, M.V. (2004).
\newblock Tolerance of pollination networks to species extinctions.
\newblock \emph{Proc. R. Soc. Lond. B}, 271, 2605--2611.

\bibitem[{Menz \emph{et~al.}(2010)Menz, Phillips, Winfree, Kremen, Aizen,
  Johnson \& Dixon}]{menz-2010-4}
\hspace{0em}\\Menz, M., Phillips, R., Winfree, R., Kremen, C., Aizen, M.,
  Johnson, S. \& Dixon, K. (2010).
\newblock Reconnecting plants and pollinators: challenges in the restoration of
  pollination mutualisms.
\newblock \emph{Trends Plant Sci.}, 16, 4--12.

\bibitem[{M'Gonigle \emph{et~al.}(2015)M'Gonigle, Ponisio, Cutler \&
  Kremen}]{mgonigle-2015-x}
\hspace{0em}\\M'Gonigle, L., Ponisio, L., Cutler, K. \& Kremen, C. (2015).
\newblock Habitat restoration promotes pollinator persistence and colonization
  in intensively-managed agriculture.
\newblock \emph{Ecol. Appl.}, 25, 1557--1565.

\bibitem[{M'Gonigle \emph{et~al.}(2016)M'Gonigle, Williams, Lonsdorf \&
  Kremen}]{mgonigle2016tool}
\hspace{0em}\\M'Gonigle, L.K., Williams, N.M., Lonsdorf, E. \& Kremen, C.
  (2016).
\newblock A tool for selecting plants when restoring habitat for pollinators.
\newblock \emph{Conserv. Lett.}

\bibitem[{Morandin \& Kremen(2013)}]{morandin-2013-829}
\hspace{0em}\\Morandin, L. \& Kremen, C. (2013).
\newblock Hedgerow restoration promotes pollinator populations and exports
  native bees to adjacent fields.
\newblock \emph{Ecol. Appl.}, 23, 829--839.

\bibitem[{Newman \& Girvan(2004)}]{Newman2004}
\hspace{0em}\\Newman, M.E.J. \& Girvan, M. (2004).
\newblock Finding and evaluating community structure in networks.
\newblock \emph{Phys. Rev. E}, 69, 026113.

\bibitem[{Nuismer \emph{et~al.}(2013)Nuismer, Jordano \&
  Bascompte}]{Nuismer2013}
\hspace{0em}\\Nuismer, S.L., Jordano, P. \& Bascompte, J. (2013).
\newblock Coevolution and the architecture of mutualistic networks.
\newblock \emph{Evolution}, 67, 338--354.

\bibitem[{Olesen \emph{et~al.}(2007)Olesen, Bascompte, Dupont \&
  Jordano}]{Olesen2007}
\hspace{0em}\\Olesen, J.M., Bascompte, J., Dupont, Y. \& Jordano, P. (2007).
\newblock The modularity of pollination networks.
\newblock \emph{Proc. Natl. Acad. Sci. USA}, 104, 19891--19896.

\bibitem[{Olesen \emph{et~al.}(2008)Olesen, Bascompte, Elberling \&
  Jordano}]{Olesen2008}
\hspace{0em}\\Olesen, J.M., Bascompte, J., Elberling, H. \& Jordano, P. (2008).
\newblock Temporal dynamics in a pollination network.
\newblock \emph{Ecology}, 89, 1573.

\bibitem[{Ollerton \emph{et~al.}(2011)Ollerton, Winfree \&
  Tarrant}]{ollerton-2011-321}
\hspace{0em}\\Ollerton, J., Winfree, R. \& Tarrant, S. (2011).
\newblock How many flowering plants are pollinated by animals?
\newblock \emph{Oikos}, 120, 321--326.

\bibitem[{Parker(1997)}]{parker1997scale}
\hspace{0em}\\Parker, V.T. (1997).
\newblock The scale of successional models and restoration objectives.
\newblock \emph{Restor. Ecol.}, 5, 301--306.

\bibitem[{Peel \& Clauset(2014)}]{peel2014detecting}
\hspace{0em}\\Peel, L. \& Clauset, A. (2014).
\newblock Detecting change points in the large-scale structure of evolving
  networks.
\newblock \emph{arXiv preprint arXiv:1403.0989}.

\bibitem[{Petanidou \emph{et~al.}(2008)Petanidou, Kallimanis, Tzanopoulos,
  Sgardelis \& Pantis}]{petanidou-2008-564}
\hspace{0em}\\Petanidou, T., Kallimanis, S., Tzanopoulos, J., Sgardelis, S. \&
  Pantis, J. (2008).
\newblock Long-term observation of a pollination network: fluctuation in
  species and interactions, relative invariance of network structure and
  implications for estimates of specialization.
\newblock \emph{Ecol. Lett.}, 11, 564--575.

\bibitem[{Ponisio \emph{et~al.}(2016)Ponisio, M'Gonigle \&
  Kremen}]{ponisio2015farm}
\hspace{0em}\\Ponisio, L.C., M'Gonigle, L.K. \& Kremen, C. (2016).
\newblock On-farm habitat restoration counters biotic homogenization in
  intensively managed agriculture.
\newblock \emph{Global Change Biol.}, 22, 704--715.

\bibitem[{Pyke(1984)}]{pyke1984optimal}
\hspace{0em}\\Pyke, G.H. (1984).
\newblock Optimal foraging theory: a critical review.
\newblock \emph{Annu. Rev. Ecol. Evol. Syst.}, pp. 523--575.

\bibitem[{{R Core Team}(2015)}]{R}
\hspace{0em}\\{R Core Team} (2015).
\newblock \emph{R: A Language and Environment for Statistical Computing}.
\newblock R Foundation for Statistical Computing, Vienna, Austria.

\bibitem[{Ramos-Jiliberto \emph{et~al.}(2012)Ramos-Jiliberto, Valdovinos,
  Moisset~de Espan{\'e}s \& Flores}]{ramos2012topological}
\hspace{0em}\\Ramos-Jiliberto, R., Valdovinos, F.S., Moisset~de Espan{\'e}s, P.
  \& Flores, J.D. (2012).
\newblock Topological plasticity increases robustness of mutualistic networks.
\newblock \emph{J. Anim. Ecol.}, 81, 896--904.

\bibitem[{Rezende \emph{et~al.}(2007)Rezende, Lavabre, Guimar{\~a}es, Jordano
  \& Bascompte}]{Rezende2007}
\hspace{0em}\\Rezende, E.L., Lavabre, J.E., Guimar{\~a}es, P.R., Jordano, P. \&
  Bascompte, J. (2007).
\newblock Non-random coextinctions in phylogenetically structured mutualistic
  networks.
\newblock \emph{Nature}, 448, 925--928.

\bibitem[{Santamar{\'\i}a \& Rodr{\'\i}guez-Giron{\'e}s(2007)}]{Santamaria2007}
\hspace{0em}\\Santamar{\'\i}a, L. \& Rodr{\'\i}guez-Giron{\'e}s, M.A. (2007).
\newblock Linkage rules for plant--pollinator networks: Trait complementarity
  or exploitation barriers?
\newblock \emph{PLoS Biol.}, 5, e31.

\bibitem[{Stang \emph{et~al.}(2009)Stang, Klinkhamer, Waser, Stang \& van
  der}]{Stang2009}
\hspace{0em}\\Stang, M., Klinkhamer, P., Waser, N.M., Stang, I. \& van der,
  M.E. (2009).
\newblock Size-specific interaction patterns and size matching in a
  plant-pollinator interaction web.
\newblock \emph{Ann. Bot.}, 103.

\bibitem[{Stang \emph{et~al.}(2006)Stang, Klinkhamer \& Van
  Der~Meijden}]{Stang2006}
\hspace{0em}\\Stang, M., Klinkhamer, P.G. \& Van Der~Meijden, E. (2006).
\newblock Size constraints and flower abundance determine the number of
  interactions in a plant--flower visitor web.
\newblock \emph{Oikos}, 112, 111--121.

\bibitem[{Th{\'e}bault \& Fontaine(2010)}]{Thebault2010}
\hspace{0em}\\Th{\'e}bault, E. \& Fontaine, C. (2010).
\newblock Stability of ecological communities and the architecture of
  mutualistic and trophic networks.
\newblock \emph{Science}, 329, 853--856.

\bibitem[{Valdovinos \emph{et~al.}(2013)Valdovinos, Moisset~de Espan{\'e}s,
  Flores \& Ramos-Jiliberto}]{valdovinos2013adaptive}
\hspace{0em}\\Valdovinos, F.S., Moisset~de Espan{\'e}s, P., Flores, J.D. \&
  Ramos-Jiliberto, R. (2013).
\newblock Adaptive foraging allows the maintenance of biodiversity of
  pollination networks.
\newblock \emph{Oikos}, 122, 907--917.

\bibitem[{Valdovinos \emph{et~al.}(2010)Valdovinos, Ramos-Jiliberto,
  Garay-Narv{\'a}ez, Urbani \& Dunne}]{valdovinos2010consequences}
\hspace{0em}\\Valdovinos, F.S., Ramos-Jiliberto, R., Garay-Narv{\'a}ez, L.,
  Urbani, P. \& Dunne, J.A. (2010).
\newblock Consequences of adaptive behaviour for the structure and dynamics of
  food webs.
\newblock \emph{Ecol. Lett.}, 13, 1546--1559.

\bibitem[{V\'azquez(2005)}]{Vazquez2005b}
\hspace{0em}\\V\'azquez, D.P. (2005).
\newblock Degree distribution in plant-animal mutualistic networks: forbidden
  links or random interactions?
\newblock \emph{Oikos}, 108, 421--426.

\bibitem[{V\'azquez \emph{et~al.}(2009)V\'azquez, Bl{\"u}thgen, Cagnolo \&
  Chacoff}]{Vazquez2009}
\hspace{0em}\\V\'azquez, D.P., Bl{\"u}thgen, N., Cagnolo, L. \& Chacoff, N.P.
  (2009).
\newblock Uniting pattern and process in plant-animal mutualistic networks: a
  review.
\newblock \emph{Ann. of Bot.}, 103, 1445--1457.

\bibitem[{V{\'a}zquez \emph{et~al.}(2009)V{\'a}zquez, Chacoff \&
  Cagnolo}]{Vazquez2009evaluating}
\hspace{0em}\\V{\'a}zquez, D.P., Chacoff, N.P. \& Cagnolo, L. (2009).
\newblock Evaluating multiple determinants of the structure of plant-animal
  mutualistic networks.
\newblock \emph{Ecology}, 90, 2039--2046.

\bibitem[{Waser \emph{et~al.}(1996)Waser, Chittka, Price, Williams \&
  Ollerton}]{Waser1996}
\hspace{0em}\\Waser, N.M., Chittka, L., Price, M.V., Williams, N.M. \&
  Ollerton, J. (1996).
\newblock Generalization in pollination systems, and why it matters.
\newblock \emph{Ecology}, 77, 1043.

\end{thebibliography}

\clearpage
\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{figures/scheme.pdf}
  \caption{Diagram illustrating the analysis to examine the temporal
    turnover of interactions weighted based on their similarity. A, B
    and C are animal species, and Ps are plant species. The dendrogram
    depicts the interaction similarity across years based on the
    number of shared constituent species.}
  \label{fig:methods}
\end{figure}
\clearpage

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/networksv1.pdf}
  \caption{Assembling hedgerow networks had more changing points
    (vertical red lines) than non-assembling hedgerows and weedy field
    margins (a representative sample of non-assembling sites are
    depicted here). In each network, plants and pollinators are
    represented by green and yellow circles, respectively, weighted by
    their degree. Each species has a consistent position in the
    perimeter of the network across years. Asterisks indicate the year
    the hedgerow was planted. Before that, the sites were weedy field
    margins.}
  \label{fig:changePoints}
\end{figure}
\clearpage

\begin{sidewaysfigure}
  \centering
  \includegraphics[width=.95\textwidth]{figures/communitiesSkanky3.pdf}
  \caption{The species module membership between network changing
    points. Two representative assembling hedgerows are depicted. In
    the top panel, species are grouped by module. The bottom panels
    visualize the flow of species to different modules between change
    points. }
  \label{fig:changePoints2}
\end{sidewaysfigure}
\clearpage

\begin{figure}
  \centering
  \includegraphics[width=.8\textwidth]{figures/occ_degree.jpeg}
  \caption{The variation coefficient of network position, as
    represented by closeness, plotted against pollinator persistence
    and degree. Persistence and degree were positively related to
    network position variability in pollinators, but unrelated in
    plants. Points represent means for each species across sites. The
    solid line indicates the mean slope estimate and the dashed lines
    are the $95\%$ confidence intervals around the estimate. }
  \label{fig:cv}
\end{figure}
\clearpage

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/turnover_panels.jpeg}
  \caption{Species interaction and weighted interactions turnover
    of plant-pollinator networks at non-assembling field margins
    sites, assembling hedgerows, and non-assembling, mature
    hedgerows. Rates of species and interaction turnover were similar
    between site types, though mature hedgerows has marginally
    significantly less pollinator turnover.  However, when
    interactions were weighted by their similarity, both hedgerow
    types had higher turnover than unrestored field margins. Boxplots
    represent medians (black horizontal line) first and third
    quartiles (box perimeter) and extremes (whiskers). }
  \label{fig:beta}
\end{figure}
\clearpage

\begin{figure}
  \centering
  \includegraphics[width=1\textwidth]{figures/baci.jpeg}
  \caption{Plant richness and pollinator richness increased as the
    networks assembled. Nestedness, modularity and specialization did
    not change predictable across years, while connectance
    decreased. The nestedness, modularity and specialization scores
    represent $z$-scores. Scores greater than $\sim 2$ or less than
    $\sim -2$ are significantly more or less structured than randomly
    assembled networks. Points are the metric value for each site at
    each year of assembly. The solid line indicates the mean slope
    estimate and the dashed lines are the $95\%$ confidence intervals
    around the estimate.}
  \label{fig:baci}
\end{figure}
\clearpage

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End: